---
title: "ACC By Behavior"
author: "Max Czapanskiy"
date: "January 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(RSQLite)
library(data.table)
library(foreach)
library(lubridate)
library(ggplot2)

# dplyr connections
MHI_db <- src_sqlite('MHI_GPS_TDR.sqlite')
metadata_db <- tbl(MHI_db, 'Metadata')
tracks_db <- tbl(MHI_db, sql('SELECT * FROM TrackView')) %>%
  filter(TripFlag == 0,
         TripEndUTC - TripStartUTC > 60 * 30) 
trips_db <- tbl(MHI_db, sql('SELECT * FROM Trips')) %>%
  filter(TripFlag == 0,
         Duration > .5)
tidytracks_db <- tbl(MHI_db, 'TidyTracks')
dives_db <- tbl(MHI_db, sql('SELECT * FROM DiveView'))

POSIX.origin = ymd('1970-01-01', tz = 'UTC')
```

## Acceleration Analysis
See Acceleration Calibration report for details on calculating overall dynamic body acceleration (ODBA).
See Behavior Classification report for details on splitting GPS tracks into rest, transit, and area-restricted search (ARS).

This code joins acceleration data with behavioral classifications and writes the results to a database table. 

### Time Lags
Because behavioral classification uses rediscretized tracks and acceleration readings are sampled at a different rate than GPS, there are lags between acceleration, behavior, and "true" locations. So in addition to annotating acceleration with behavioral state, we also note two time lags. The acceleration lag is the interval between an acceleration reading and the *latest* behavior before it (i.e. rolling forward from behavior). The behavior lag is the interval between a behavior and the *closest* GPS point (i.e rolling nearest from GPS). Only acceleration-behavior tuples with lags below a threshold (e.g. 5 minutes) will be retained for later analyses.
```{r join, cache=TRUE}
# List of acceleration CSVs (output from Acceleration Calibration report)
calib.files <- dir('E-Obs/Calibration/', 
                  pattern = 'RFBO[0-9]{2}_[0-9]{4}.csv', 
                  full.names = TRUE)

# Iterating through acceleration CSVs...
foreach(acc.file = calib.files,
        i = seq(calib.files),
        .combine = rbind) %do% {
          
  # Acceleration data read from file joined with metadata
  acc.data <- read.csv(acc.file,
                       stringsAsFactors = FALSE) %>%
    mutate(ACCTimestampUTC = ymd_hms(timestamp, tz = 'UTC'),
           DummyTimestampUTC = ACCTimestampUTC)
  metadata <- metadata_db %>%
    filter(ACC_ID == acc.data$TagID[1]) %>%
    select(TagID = ACC_ID,
           DeployID) %>%
    collect
  acc.data$DeployID = metadata$DeployID
  acc.dt <- data.table(acc.data, key = 'DummyTimestampUTC')
  
  # Behavior data read from database
  behavior.dt <- tidytracks_db %>%
    filter(DeployID == metadata$DeployID) %>%
    select(DeployID,
           TripID,
           TimestampUTC,
           Longitude,
           Latitude,
           Behavior = behavior) %>%
    collect(n = Inf) %>%
    mutate(BehaviorTimestampUTC = as.POSIXct(TimestampUTC, origin = POSIX.origin, tz = 'UTC'),
           DummyTimestampUTC = BehaviorTimestampUTC) %>%
    data.table(key = 'DummyTimestampUTC')
  
  # Join acceleration with behavior data
  acc.behavior0 <- behavior.dt[acc.dt, roll = TRUE] %>%
    select(DeployID, 
           TripID, 
           TagID, 
           BehaviorTimestampUTC, 
           Behavior,
           ACCTimestampUTC, 
           Longitude,
           Latitude,
           ends_with('ODBA'),
           DummyTimestampUTC)
  setkey(acc.behavior0, DummyTimestampUTC)
  
  # "True" GPS readings read from database
  tracks.dt <- tracks_db %>%
    filter(DeployID == metadata$DeployID) %>%
    select(DeployID, 
           TimestampUTC) %>%
    collect(n = Inf) %>%
    mutate(TrueTimestampUTC = as.POSIXct(TimestampUTC, origin = POSIX.origin, tz = 'UTC'),
           DummyTimestampUTC = TrueTimestampUTC) %>%
    data.table(key = 'DummyTimestampUTC')
  
  # Join "true" GPS readings with acceleration/behavior and calculate lags
  acc.behavior <- tracks.dt[acc.behavior0, roll = 'nearest'] %>% 
    select(DeployID, 
           TripID, 
           TagID, 
           Longitude,
           Latitude,
           BehaviorTimestampUTC, 
           Behavior, 
           ACCTimestampUTC, 
           ends_with('ODBA'), 
           TrueTimestampUTC) %>%
    mutate(BehaviorLag = BehaviorTimestampUTC - TrueTimestampUTC,
           ACCLag = ACCTimestampUTC - TrueTimestampUTC) %>%
    na.omit
  
  # Write to database
  dbWriteTable(MHI_db$con, 'ACC_Behavior', acc.behavior, overwrite = i == 1, append = i > 1)
  
  # Summarize effort
  acc.behavior %>%
    summarize(DeployID = first(DeployID),
              TagID = first(TagID),
              Trips = n_distinct(TripID),
              N = n())
}
```

## Visualize results
```{r plot}
tbl(MHI_db, 'ACC_Behavior') %>%
  # Only include behaviors within five minutes of a recorded point
  filter(abs(BehaviorLag) < 5 * 60) %>%
  collect(n=Inf) %>%
  mutate(Behavior = factor(Behavior, levels = c('rest', 'transit', 'ars'), labels = c('Rest', 'Transit', 'ARS'))) %>%
  ggplot(aes(x = Behavior, y = meanODBA)) +
  geom_violin() +
  theme_bw() +
  labs(title = 'Distribution of ODBA Across Behaviors',
       subtitle = 'Tri-axial acceleration measured in bursts every thirty seconds at 20hz for 4.8s',
       y = 'Mean ODBA\n(Per burst)')
```
