---
title: "Acceleration Calibration"
author: "Max Czapanskiy"
date: "January 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(foreach)
library(iterators)
library(doParallel)
library(tidyr)
library(lubridate)
library(ggplot2)
library(move)
library(RSQLite)
library(httr)

select <- dplyr::select

httr::set_config( config( ssl_verifypeer = 0L ) )
curl <- movebankLogin(readLines('movebank.txt')[1], readLines('movebank.txt')[2])
MovebankID <- 169686018

cl <- makeCluster(detectCores())
registerDoParallel(cl)
```

## Acceleration Calibration
Static vs dynamic
Convert readings to m/s^2

Field	Meaning
TagID ID number of e-obs tag
A	    -z orientation start time (see pages 4-5 of e-obs GPS-Acceleration-Tags manual)
B	    +z orientation start time
C	    -x orientation start time
D	    +x orientation start time
E	    +y orientation start time
F	    -y orientation start time

Fit a line to (x1, -g), (x2, +g) [where x1 is C and x2 and D] to get acceleration in x-axis. Similarly for y and z.
a = mr + b
where a is acceleration in m/s^2 and r is the raw measurement.
-g = m * A + b

```{r calib, warning=FALSE}
calib.meta <- read.csv('E-Obs/Calibration/e-obs_calibration.csv', stringsAsFactors = FALSE) %>%
  mutate_at(.cols = vars(A:F), .funs = funs(as.POSIXct(. , tz = 'US/Pacific')))

# Assumes tagxxxx_acc.txt has ACC in columns
# i.e. decoder.exe -> 6 -> c
calib.const <- foreach(tag = dir('E-Obs/Calibration', pattern = '.*_acc.txt'), .combine = rbind) %do% {
  tagid <- substr(tag, 4, 7) %>% as.integer
  calib.data <- read.csv(sprintf('E-Obs/Calibration/%s', tag),
                         stringsAsFactors = FALSE,
                         col.names = c('jDate', 'date', 'day', 'time', 'X', 'Y', 'Z')) %>%
    mutate(Tag.ID = tagid,
           date.time = paste(date, time) %>% dmy_hms(tz = 'UTC')) 
  
  mean.acc <- foreach(position = LETTERS[1:6],
                      start.time = filter(calib.meta, Tag.ID == tagid) %>% select(A:F),
                      .combine = rbind) %do% {
    begin <- start.time + minutes(1)
    end <- start.time + minutes(9)
    # Calibration metadata is in UTC-7 but calibration data is in UTC+0 so you'll get an inconsistent tzone warning
    filter(calib.data, date.time > begin, date.time < end) %>%
      summarize_at(.cols = vars(X:Z), .funs = funs(mean)) %>%
      mutate(Position = position,
             Begin = begin,
             End = end)
  } %>%
    transmute(Position,
              Begin,
              End,
              Measurement = c(Z[1], Z[2], X[3], X[4], Y[5], Y[6]),
              Orientation = c('-z', '+z', '-x', '+x', '+y', '-y'),
              TagID = tagid)
  
  plot.data <- calib.data %>%
    select(date.time, X:Z) %>%
    gather(key = dimension, value = measurement, X:Z) %>% 
    arrange(date.time, dimension) 
  
  p <- ggplot(plot.data,
         aes(x = date.time, y = measurement, color = dimension)) +
    geom_segment(aes(x = Begin, xend = End,
                     y = Measurement, yend = Measurement,
                     color = Orientation),
                 data = mean.acc,
                 size = 2,
                 show.legend = FALSE) +
    geom_line() +
    labs(title = sprintf('ACC Calibration: Tag %i', tagid),
         x = 'Time',
         y = 'Raw Reading')
  
  ggsave(sprintf('E-Obs/Calibration/Plots/%i.png', tagid), p, height = 7.5, width = 10, units = 'in')
  
  g <- 9.820 #m/s^2
  x1 = mean.acc$Measurement[3]
  x2 = mean.acc$Measurement[4]
  y1 = mean.acc$Measurement[6]
  y2 = mean.acc$Measurement[5]
  z1 = mean.acc$Measurement[1]
  z2 = mean.acc$Measurement[2]
  data.frame(TagID = tagid) %>%
    mutate(mX = -2 * g / (x1 - x2),
           bX = -g - mX * x1,
           mY = -2 * g / (y1 - y2),
           bY = -g - mY * y1,
           mZ = -2 * g / (z1 - z2),
           bZ = -g - mZ * z1)
}
```

```{r analyze}
getMovebankAccData <- function(study = MovebankID, login = curl, animalName = 'RFBO01') {
  d <- getMovebank("individual", login=login, study_id=study, attributes=c('id','local_identifier'))
  animalName <- d[as.character(d$local_identifier) %in% animalName,'id']
  
  individual <- getMovebank("individual", login=login, study_id=study, id=animalName, attributes=c("id", "local_identifier")) %>%
    rename(individual_id = id,
           animal_name = local_identifier)
  deployment <- getMovebank("deployment", login=login, study_id=study, attributes=c("individual_id","tag_id","id"), individual_id=animalName) %>%
    rename(deployment_id = id)
  tags <- getMovebank("tag", login=login, study_id=study, attributes=c("id", "local_identifier")) %>%
    rename(tag_id = id,
           tag_name = local_identifier)
  
  idData <- merge(individual, deployment) %>%
    merge(tags)
  
  accSen <- 2365683
  attribs <- c("timestamp", "eobs_acceleration_axes", "eobs_acceleration_sampling_frequency_per_axis", "eobs_accelerations_raw", "sensor_type_id", "deployment_id", "event_id")
  
  trackDF <- getMovebank("event", login = login, study_id = study, attributes = attribs , deployment_id = idData$deployment, sensor_type_id = accSen)

  if(nrow(trackDF)==0){
    stop('No records found for this individual/study combination')
  }
  
  trackDF$timestamp <- as.POSIXct(strptime(as.character(trackDF$timestamp), format = "%Y-%m-%d %H:%M:%OS",tz="UTC"), tz="UTC")
  
  if(any(tapply(trackDF$sensor_type_id, trackDF$deployment_id, length)!=1)){# data comes in ordered by sensor but needs to be ordered by timestamp
    trackDF <- trackDF[ with(trackDF, order(trackDF$deployment_id, timestamp)) , ]  
  }
  
  trackDF$TagID <- idData$tag_name
  
  trackDF
}

AccSensorID <- 2365683
tags <- getMovebank("tag", login=curl, study_id=MovebankID, attributes=c("id", "local_identifier")) %>%
  rename(tag_id = id,
         tag_name = local_identifier)
animals <- getMovebankAnimals(MovebankID, curl) %>%
  filter(sensor_type_id == AccSensorID) %>%
  merge(tags)

foreach(animal_dep = animals$animalName_deployment, tag = animals$tag_name) %do% {
  animal <- substr(animal_dep, 1, 6)
  const <- filter(calib.const, TagID == tag)
  m <- select(const, starts_with('m')) %>% as.numeric
  b <- select(const, starts_with('b')) %>% as.numeric
  
  print(sprintf('Fetching acceleration for %s (tag %i)...', animal, tag))
  eobs.data <- getMovebankAccData(animalName = animal) %>% 
    transmute(TagID, 
              timestamp, 
              ACC = as.character(eobs_accelerations_raw))
  
  print('Calculating ODBA...')
  odba.data <- 
    foreach(row = iter(eobs.data, by = 'row'),
            .combine = bind_rows,
            .multicombine = TRUE,
            .packages = 'dplyr') %dopar%
    {
      raw.acc <- strsplit(row$ACC, ' ') %>%
        unlist(use.names = FALSE) %>%
        as.numeric %>%
        matrix(ncol = 3, byrow = TRUE)
      colnames(raw.acc) <- c('X0', 'Y0', 'Z0')
      
      acc <- data.frame(raw.acc) %>%
        mutate(totX = m[1] * X0 + b[1],
               totY = m[2] * Y0 + b[2],
               totZ = m[3] * Z0 + b[3],
               staticX = mean(totX),
               staticY = mean(totY),
               staticZ = mean(totZ),
               dynX = totX - staticX,
               dynY = totY - staticY,
               dynZ = totZ - staticZ)
      
      odba <- with(acc, sqrt(dynX^2 + dynY^2 + dynZ^2))
      
      row %>%
        select(TagID, timestamp) %>%
        mutate(meanODBA = mean(odba),
               sdODBA = sd(odba),
               minODBA = min(odba),
               maxODBA = max(odba))
    }
  
  print('Writing to file...')
  write.csv(odba.data, sprintf('E-Obs/Calibration/%s_%i.csv', animal, tag), row.names = FALSE)
}
```
