---
title: "Behavioral Classification"
author: "Max Czapanskiy"
date: "December 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(ggplot2)
library(proj4)
library(lubridate)
library(dplyr)
library(geosphere)
library(mosaic)
library(adehabitatLT)
library(foreach)
library(doParallel)
library(iterators)
library(RSQLite)
library(DBI)
library(tcltk)
select <- dplyr::select

# dplyr connections
MHI_db <- src_sqlite('MHI_GPS_TDR.sqlite')
metadata_db <- tbl(MHI_db, 'Metadata')
tracks_db <- tbl(MHI_db, sql('SELECT * FROM TrackView')) %>%
  filter(TripEndUTC - TripStartUTC > 60 * 30) 
trips_db <- tbl(MHI_db, sql('SELECT * FROM Trips')) %>%
  filter(Duration > .5)
tidytracks_db <- tbl(MHI_db, 'TidyTracks')
dives_db <- tbl(MHI_db, sql('SELECT * FROM DiveView'))

POSIX.origin = ymd('1970-01-01', tz = 'UTC')
RST_tracker <- 'RST/RST_Output/RST_tracker.csv'

cl <- makeCluster(5)
registerDoParallel(cl)
```

```{r setupx32, include = FALSE}
dyn.load('RST/RST_residenceRadii.dll')
source('RST/RST_functions_all.R')
```

## Behavioral Classification

This script classifies tracks using three behavioral states: transit, area-restricted search (ARS), and rest. Classification takes place in two phases. First, resting points are identified using a speed threshold (e.g. rest = speed < 1 m/s). The remaining points are divided between transit and ARS using the Residence in Space and Time (RST) algorithm (Torres et al. n.d.). For each point along the track, RST calculates a residence time (RT, time it takes to cross a circle of radius *r*) and residence distance (RD, distance traversed while crossing a circle of radius *r*), where *r* is defined as mean transit speed \* sampling interval / 2. Normalizing these residence values (nRT, nRD) and taking the difference provides the basis for classification (*res* = nRD - nRT). Thus, points with positive *res* are more distance-intensive than time-intensive and are classified as ARS. Conversely, points with negative *res* are more time-intensive than distance-intensive and are added to the resting classification. Points where *res* = 0 have high speeds and low tortuosity and re classified as transit.

 * Only trips with durations greater than 30 minutes and TripFlag = 0 (i.e. no issues) are considered for behavioral classification.
 * The RST library only works with 32-bit R. Many systems use 64-bit R by default so watch out!
 
### Rediscretize and calculate speed/tortuostiy
Set R to 64-bit for reliability

 
```{r rediscretize, eval=FALSE}
#rediscretized <- foreach(species = c('BRBO', 'LAAL', 'RFBO', 'RTTR', 'WTSH'), .combine = rbind) %do% {
rediscretized <- foreach(species = 'RFBO', .combine = rbind) %do% {
  trips <- trips_db %>%
    filter(Species == species) %>%
    collect
  
  pb <- tkProgressBar(title = "Tidying tracks", label = "", min = 0, max = nrow(trips), initial = 0, width = 300)

  foreach(did = trips$DeployID, tn = trips$TripNumber, i = seq(trips$DeployID), .combine = rbind) %do% {
    setTkProgressBar(pb, i, label = sprintf('%s DID:%i TN:%i', species, did, tn))
    
    # Load trip track
    raw.track <- filter(tracks_db, 
                        DeployID == did,
                        TripNumber == tn) %>%
      arrange(TimestampUTC) %>%
      collect %>%
      transmute(band = as.numeric(sprintf('%i%05i', DeployID, TripNumber)),
                lat = Latitude,
                lon = Longitude,
                datetime = as.POSIXct(TimestampUTC, origin = POSIX.origin, tz = 'UTC'))
    
    # Create grid x- and y- coordinates from the longitude/latitude coordinates
    proj.string <- '+proj=lcc +lat_1=20 +lat_2=40 +lon_0=160 +lat_0=-10'
    lambert <- proj4::project(raw.track[, c('lon', 'lat')],
                              proj = proj.string)
    raw.track$x <- lambert$x
    raw.track$y <- lambert$y
    
    # Rediscretize track to 3 min (for now; will vary by deployment) 
    time_step <- 2 * 60
    # Yields a ltraj (adehabitatLT)
    redis.track <- as.ltraj(raw.track[,c('x','y')],
                            date = raw.track$datetime,
                            id = raw.track$band) %>%
      redisltraj(u = time_step,
                 samplex0 = TRUE,
                 type = 'time')
    
    # Calculate speed and tortuosity
    track <- redis.track %>%
      ld %>% # Convert ltraj to data.frame
      mutate(dd = dist,
             speed = dd / dt,
             cumdd = cumsum(dd),
             tortuosity = (lead(cumdd, 4) - lag(cumdd, 6)) / 
               sqrt((lead(x, 5) - lag(x, 5))^2 + (lead(y, 5) - lag(y, 5))^2)) %>%
      transmute(TripID = burst,
                TimestampUTC = date,
                x,
                y,
                DeployID = did,
                TripNumber = tn,
                Species = species,
                Year = year(date[1]),
                dd,
                dt,
                speed,
                tortuosity)
    
    # Invert projection to get longitude/latitude in WGS84
    wgs84 <- proj4::project(track[, c('x', 'y')],
                            proj = proj.string,
                            inverse = TRUE)
    track$Longitude <- wgs84$x
    track$Latitude = wgs84$y
    
    # Trip QAQC flags
    TripQAQC <- trips %>%
      filter(DeployID == did,
             TripNumber == tn) %>%
      select(TripFlag:BehaviorOK) %>%
      slice(1) 

    track <- mutate(track, 
                    nRT = NA, nRD = NA, residual = NA, behavior = NA) %>%
      merge(TripQAQC)
  
    # Write to DB
    # INTEGERs are converted to REALs - what gives?
    dbWriteTable(MHI_db$con, 'TidyTracks', track, append = TRUE)
    
    data.frame(species = species, deployid = did, tripnumber = tn)
  } 
}
```

### Classify by RST
Set R to 32-bit for compatibility with RST DLL

```{r classify, eval=FALSE}
#foreach(species = c('BRBO', 'LAAL', 'RFBO', 'RTTR', 'WTSH'), .combine = rbind) %do% {
classified <- foreach(species = 'RFBO', .combine = rbind) %do% {  
  time_step <- 2 * 60
  
  # Mean transit speed
  median.tortuosity.df <- filter(tidytracks_db,
                                 Species == species,
                                 speed > 1,
                                 TripFlag == 0) %>%
    select(tortuosity) %>%
    collect(n = Inf)
  median.tortuosity <- median(median.tortuosity.df$tortuosity, na.rm = TRUE)
  mean.transit.speed.df <- filter(tidytracks_db,
                                  Species == species,
                                  speed > 1,
                                  TripFlag == 0,
                                  tortuosity < median.tortuosity) %>%
    select(speed) %>%
    collect(n = Inf)
  mean.transit.speed <- mean(mean.transit.speed.df$speed, na.rm = TRUE)
  
  radius <- mean.transit.speed * time_step / 2 # in meters
  threshold <- 0
  
  # trips <- trips_db %>%
  #   filter(Species == species,
  #          Year >= 2013,
  #          BehaviorOK == 'y') %>%
  #   collect
  trips <- trips_db %>%
    filter(Species == species,
           Year == 2016) %>%
    collect
  
  time_units <- 'secs'
  
  pb <- tkProgressBar(title = "Classifying tracks", label = "", min = 0, max = nrow(trips), initial = 0, width = 300)
  foreach(tid = trips$TripID, i = seq(trips$TripID), .combine = rbind) %do% {
    setTkProgressBar(pb, i, label = sprintf('%s TID:%i', species, tid))
    # Load species' tracks
    tracks <- filter(tidytracks_db, 
                     TripID == tid) %>%
      arrange(TimestampUTC) %>%
      collect(n = Inf) %>%
      mutate(datetime = as.POSIXct(TimestampUTC, origin = POSIX.origin, tz = 'UTC'),
             time_diff = as.numeric(datetime - datetime[1], units = time_units))
    
    if(with(tracks, as.numeric(last(tracks$datetime) - first(tracks$datetime), units = 'hours') < .5)) return(NULL)
    
    rst <- with(tracks, residenceCalc(x, y, time_diff, radius, threshold)) %>%
      select(nRT = 4, nRD = 5, residual = 6)
    
    tracks.rst <- tracks %>%
      select(-datetime, -time_diff, -nRT, -nRD, -residual, -behavior) %>%
      cbind(rst) %>%
      mutate(behavior = derivedFactor(
        rest = (residual < 0 | speed < 1),
        ars = residual > 0,
        transit = residual == 0,
        .method = 'first'))
    
    out.file <- sprintf('RST/RST_Output/%s.csv', tid)
    write.csv(tracks.rst, out.file, row.names = FALSE)
    data.frame(rst.file = out.file)
  }
}

write.csv(classified, RST_tracker, row.names = FALSE)
```

### Write classifications to DB
Switch back to 64-bit

```{r classify2, eval=FALSE}
rst.files <- read.csv(RST_tracker, stringsAsFactors = FALSE)$rst.file
pb <- tkProgressBar(title = "Uploading classifications", label = "", min = 0, max = length(rst.files), initial = 0, width = 300)

foreach(csv = rst.files, i = seq(rst.files)) %do% {
  rst.track <- read.csv(csv, stringsAsFactors = FALSE)
  tid <- rst.track$TripID[1]
  setTkProgressBar(pb, i, label = sprintf('TID:%i', tid))
 
  rst.track <- select_(rst.track, .dots = colnames(tidytracks_db))
  q <- dbSendStatement(MHI_db$con, sprintf('DELETE FROM TidyTracks WHERE TripID = %i', tid))
  dbClearResult(q)
  dbWriteTable(MHI_db$con, 'TidyTracks', rst.track, append = TRUE)  
}
```
