---
title: "ASSP CPUE GLMM"
author: "Amelia DuVall (ajduvall@uw.edu)"
date: "July 2020"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is `r paste0("v.", (Sys.Date()))`

```{r set_options, message = FALSE}
knitr::opts_chunk$set(message = FALSE)

## load packages
library(easypackages)
libraries("lme4", "here", "tidyverse", "ggplot2", "lubridate", 
          "faraway", "anchors", "readxl", "naniar", "remotes",
          "corrplot", "Hmisc")
```

see Russ Bradley et al. 2011 paper for ideas

```{r data}
## read-in data 
wx <- readRDS(here("Working", "mistnet_w_weather_v2.rds")) # 215 sessions
CPUE <- readRDS(here("Working", "cpue.RDS"))

## join with session_ID
dfr <- inner_join(wx, CPUE, by = c("session_ID", "island_code", "site_code", "lat", "long")) %>%
  replace_with_na(replace = list("tot_precip" = "NaN", "wind_u" = "NaN", "wind_v" = "NaN", 
                                 "wave_height" = "NaN", "wave_dir" = "NaN", "wave_period" = "NaN",
                                 "tot_cloud" = "NaN", "low_cloud" = "NaN", "pressure" = "NaN", 
                                 "surface_pressure" = "NaN", "sst" = "NaN", "temp_c" = "NaN",
                                 "dew_point" = "NaN", "Moonup" = "NaN", "MoonAltitude" = "NaN", 
                                 "Illu"= "NaN")) %>%
  mutate(wind_speed = sqrt(wind_u^2+wind_v^2)) %>%
  dplyr::select("session_ID", "session_year", "session_month", "island_code", "site_code", 
                "start", "end", "wind_speed", "tot_precip", "wave_height":"Illu", "CPUEstd") %>%
  filter(!is.na(CPUEstd)) %>% # 197 obs
  mutate(logCPUE = log(CPUEstd)) %>% # use log of CPUEstd
  ungroup()
  
## check CPUEstd NAs
NAs <- CPUE %>%
  filter(is.na(CPUEstd)) # 40 obs

## scale predictors
df <-  dfr %>% ## issues trying this with mutate_at
  mutate(tot_precip = scale(tot_precip),
         wind_speed = scale(wind_speed), 
         wave_height = scale(wave_height),
         wave_dir = scale(wave_dir), 
         wave_period = scale(wave_period),
         tot_cloud = scale(tot_cloud), 
         low_cloud = scale(low_cloud),
         pressure = scale(pressure), 
         surface_pressure = scale(surface_pressure),
         sst = scale(sst), 
         temp_c = scale(temp_c), 
         dew_point = scale(dew_point),
         Moonup = scale(Moonup), 
         MoonAltitude = scale(MoonAltitude), 
         Illu = scale(Illu))

## explore data
ggplot(df) +
  geom_histogram(mapping = aes(x = CPUEstd), bins = 20) +
  xlab("CPUEstd") + ylab("Frequency") +
  theme_minimal()
# Appropriate to use normal distribution?

## try logCPUE
ggplot(df) +
  geom_histogram(mapping = aes(x = logCPUE), bins = 20) +
  xlab("logCPUE") + ylab("Frequency") +
  theme_minimal()
# Looks better!

## other plots
effort <- df %>%
  group_by(session_year, site_code) %>%
  summarize(effort = n())

ggplot(effort) + 
  geom_point(mapping = aes(x = as.character(session_year), y = effort, color = site_code), 
             position = "jitter") +
  xlab("Year") + ylab("No. sessions") +
  theme_minimal()
```

## Modeling
```{r nullmodel}
## fit intercept-only model
# site and year as random effects
nmRE <- lmer(logCPUE ~ 1 + (1 | site_code) + (1 | session_year), data = df)
sumary(nmRE)
plot(nmRE)

## model diagnostics
# set plot area
par(mai = c(0.9, 0.9, 0.6, 0.1),
omi = c(0, 0, 0, 0),
mfrow = c(1,2), cex.lab = 1.2)

## qq resids
qqnorm(residuals(nmRE), main = "QQ plot (residuals)", las = 1, pch = 16)
qqline(residuals(nmRE)) # doesn't look too bad

## residuals v fitted values
plot(fitted(nmRE), residuals(nmRE), las = 1, pch = 16,
xlab = "Fitted", ylab = "Residuals",
main = "Residuals vs fitted")
abline(h=0, lty = "dashed") # also does not look too bad

# ## other models
# # island as explanation for CPUE?
# isl <- lm(CPUEstd ~ island_code, data = df)
# sumary(isl)
# plot(isl)
# 
# # try w/o random effects
# nm <- lm(logCPUE ~ 1, data = df)
# summary(nm)
```

## Predictors from ERA5 reanalysis data
https://confluence.ecmwf.int/display/CKB/ERA5%3A+data+documentation#ERA5:datadocumentation-Parameterlistings

* wind_u = '10m_u_component_of_wind' = 10 metre U wind component (m s**-1)
      # This parameter is the eastward component of the 10m wind. It is the horizontal speed of air moving towards        #  the east, at a height of ten metres above the surface of the Earth, in metres per second.
      # https://apps.ecmwf.int/codes/grib/param-db?id=165
* wind_v = '10m_v_component_of_wind' = 10 metre V wind component (m s**-1)
      # 	This parameter is the northward component of the 10m wind. It is the horizontal speed of air moving 
      # towards the north, at a height of ten metres above the surface of the Earth, in metres per second.
      # https://apps.ecmwf.int/codes/grib/param-db?id=166
* dew_point = '2m_dewpoint_temperature' = 2 metre dewpoint temperature (K)
* temp_C = '2m_temperature' =  2 metre temperature (K) 
      ## confirm this with Abram, diff code online "2t", "t2m" in era5 code from Abram
      # https://apps.ecmwf.int/codes/grib/param-db?id=167
* low_cloud = 'low_cloud_cover' = Low cloud cover (0 - 1)
* pressure = 'mean_sea_level_pressure'=  Mean sea level pressure (Pa)
* wave_dir = 'mean_wave_direction' = Mean wave direction (degrees)
* wave_period = 'mean_wave_period' = Mean wave period (s)
* sst = 'sea_surface_temperature', = Sea surface temperature (K)
* wave_height = 'significant_height_of_combined_wind_waves_and_swell'= Significant height of combined wind waves and swell (m)
* surface_pressure = 'surface_pressure' = Surface pressure (Pa)
* tot_cloud = 'total_cloud_cover' = Total cloud cover (0 - 1)
* tot_precip = 'total_precipitation' = Total precipitation (m)

## Moon data from oce package
* moonup = 1 if moon is above horizon, otherwise 0
      ## should these values be averaged then?
* moonaltitude = altitude = moon altitude, in degrees from -90 to 90.
* illu = 'illuminatedFraction' = fraction of moon’s visible disk that is illuminated

### Correlation between predictors 
```{r corr}
## check for correlation between predictors (cutoff >0.65)
(cp <- as.data.frame((round(cor(df[,10:25], use="complete.obs"), 2))))

# covs df
covs <- dfr %>%
  dplyr::select(wind_speed:Illu)

# this function flattens your data in a particular way, used below
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
  )
}

cor <- rcorr(as.matrix(covs))

# creates a new formatted df
cor_vals <- flattenCorrMatrix(cor$r, cor$P) %>% arrange(cor)

#makes a plot
corrplot(cor$r, type = "lower", order = "original", p.mat = cor$P,
         insig = "blank", sig.level = 0.01, tl.col = "black", tl.cex = .9, number.cex = .2)
```
Positive correlations are displayed in blue and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients.

Correlated predictors:
* wind_u and wind_v 
* wind_v and wave_height
* wave_period and wave_dir
* low_cloud and tot_cloud
* surface_pressure and pressure
* temp_c and sst
* dew_point and sst
* dew_point and temp_c
* moonup and moonaltitude
* moonup and illu
* moonaltitude and illu
* moonup and moonaltitude

*wing_speed = sqrt(wind_u^2+wind_v^2)*
*moon time above horizon x illu*

```{r mod}
preds <- c("tot_precip", "wind_u", "wind_v", "wave_height", "wave_dir", 
               "wave_period","tot_cloud", "low_cloud", "pressure", "surface_pressure", 
               "sst", "temp_c", "dew_point", "Moonup", "MoonAltitude", "Illu")
# Combine moon variables? Wind? Wave?
# How to deal with wind? Extract wind speed at midnight? east and north variables

top_preds <- c("tot_precip", "wind_u", "wave_height", "tot_cloud", "pressure", 
               "sst", "temp_c", "dew_point", "MoonAltitude", "Illu")

## model selection
## create data frame specifying predictors to include
predictors <- as.data.frame(matrix(c(FALSE, TRUE), 2, 10)) # 10 potential predictors

## add column names 
cov_names <- colnames(predictors) <- c("tot_precip", "wind_u", "wave_height", "tot_cloud", " pressure", 
               "sst", "temp_c", "dew_point", "MoonAltitude", "Illu")

## create set of all possible combinations
full_set <- expand.grid(predictors) # 1024 combinations

## select models with correlated predictors
ii <- which(full_set$temp_c + full_set$sst == 2 |
              full_set$dew_point + full_set$sst == 2 |
              full_set$dew_point + full_set$temp_c == 2 |
              full_set$MoonAltitude + full_set$Illu == 2)

# ii <- which(full_set$wind_u + full_set$wind_v == 2 |
#               full_set$wind_v + full_set$wave_height == 2 |
#               full_set$wave_period + full_set$wave_dir == 2 |
#               full_set$low_cloud + full_set$tot_cloud == 2 |
#               full_set$surface_pressure + full_set$pressure == 2 |
#               full_set$temp_c + full_set$sst == 2 |
#               full_set$dew_point + full_set$sst == 2 |
#               full_set$dew_point + full_set$temp_c == 2 |
#               full_set$Moonup + full_set$MoonAltitude == 2 |
#               full_set$Moonup + full_set$Illu == 2 |
#               full_set$MoonAltitude + full_set$Illu == 2 |
#               full_set$Moonup + full_set$MoonAltitude == 2)

## create reduced set of models and convert to a matrix for easier indexing
use_set <- as.matrix(full_set[-ii,]) # 384 models to fit

## number of models in set
(n_mods <- nrow(use_set)) # 384 models to fit

## create empty matrix for storing results
mod_res <- matrix(NA, n_mods, 2)
colnames(mod_res) <- c("AIC", "BIC")

## fit models and store AIC & BIC
for(i in 1:n_mods) {
  if(i == 1) {
    fmla <- "logCPUE ~ 1 + (1 | site_code) + (1 | session_year)"
  } else {
    fmla <- paste("logCPUE ~ (1 | site_code) + (1 | session_year) +", paste(cov_names[use_set[i,]], collapse = " + "))
  }
  mod_fit <- lmer(as.formula(fmla), data = df)
  mod_res[i,"AIC"] <- AIC(mod_fit)
  mod_res[i,"BIC"] <- BIC(mod_fit) 
}

## create empty matrix for storing results
delta_res <- matrix(NA, n_mods, 2)
colnames(delta_res) <- c("deltaAIC", "deltaBIC")

## convert IC to deltaIC
delta_res[,"deltaAIC"] <- mod_res[,"AIC"] - min(mod_res[,"AIC"])
delta_res[,"deltaBIC"] <- mod_res[,"BIC"] - min(mod_res[,"BIC"])
(delta_res <- round(delta_res, 2)) # round results

## create df with mod results
mp <- as.data.frame(use_set) 

allm <- as.data.frame(delta_res) %>%
  mutate(modelno = 1:384)%>%
  arrange(deltaAIC) 

## create df with top models
bestm <- allm %>%
  filter(deltaAIC <= 2) # 5 competitive models 

top <- mp %>%
  slice(97, 33, 353, 65, 321)

## final predictors
## fit full model
fm <- lm(logCPUE ~ Type + PDO + Fish + PDSI + CUTI + (1 | Plot) + (1 | Year), 
            data = SCMU, family = binomial)
summary(fm)
```

```{r bestmod, echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
bm <- lmer(logCPUE ~ (1 | site_code) + (1 | session_year) + dew_point, data = df)
summary(bm)
sumary(bm)
mm <- model.matrix(bm)

## extract predicted values and plot
preds <- predict(bm)
ggplot() + geom_histogram(mapping = aes(preds)) + theme_minimal()

## extract residuals and plot
resids <- residuals(bm)
ggplot() + geom_histogram(mapping = aes(resids)) + theme_minimal()

## extract coeffs and random effects
coef(bm)
artificial <- 1 / (1 + exp(-(0.15064))) # intercept
crevice <- 1 / (1 + exp(-(0.15064 + 0.04569))) 
shrub <- 1 / (1 + exp(-(0.15064 + 1.02356))) 
ranefyr <- ranef(bm)$Year
ranefpl <- ranef(bm)$Plot

## estimated mean success
eta <- predict(bm)

## mean response
mean(est <- 1 / (1 + exp(-eta)))
```


```{r diagnostics, echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
## Q-Q Plots
## set plot area
par(mai = c(0.9, 0.9, 0.6, 0.1),
    omi = c(0, 0, 0, 0),
    mfrow = c(1,2), cex.lab = 1.2)

# qq resids
qqnorm(resids, main = "QQ plot (residuals)", las = 1, pch = 16)
qqline(resids)

# qq Plot RE
qqnorm(unlist(ranefpl), main = "QQ plot (Plot RE)", las = 1, pch = 16)
qqline(unlist(ranefpl))

# qq Year RE
qqnorm(unlist(ranefyr), main = "QQ plot (Year RE)", las = 1, pch = 16)
qqline(unlist(ranefyr))

## plot residuals versus fitted values
yh <- fitted(bm, type = "response")
plot(yh, resids, las = 1, pch = 16,
     xlab = "Fitted", ylab = "Residuals",
     main = "Residuals vs fitted")
abline(h=0, lty = "dashed")

## Levene's test
## split residuals into 2 groups
g1 <- resids[yh <= median(yh)]
g2 <- resids[yh > median(yh)]

## Levene's test
var.test(g1, g2)
# We would reject the null hypothesis that the variances are equal. 

## Goodness of fit
## Pearson's X^2 statistic
X2 <- sum((success - yh)^2/yh)
df <- nn-5 # 3 fixed effects, 2 random effects

## likelihood ratio test
pchisq(X2, df = 1323, lower.tail = FALSE)
# large p-value, cannot reject H_0

## check for overdispersion 
c_hat <- deviance(bm)/df
# close to 1

## Likelihood ratio tests  
## run model with plot RE only
bm_plot <- glmer(Success ~ Type + (1 | Plot), data = SCMU, family = binomial)
## run model with year RE only
bm_yr <- glmer(Success ~ Type + (1 | Year), data = SCMU, family = binomial)
## run model with both REs (same as top model)

## conduct an LRT to see if the variance of the Plot RE is contributing useful info
test_1 <- 2 * (logLik(bm) - logLik(bm_yr))
pchisq(as.numeric(test_1), df = 1, lower.tail = FALSE)
# There is support for inclusion of Plot as an RE

## check for contribution of Year RE
test_2 <- 2 * (logLik(bm) - logLik(bm_plot))
pchisq(as.numeric(test_2), df = 1, lower.tail = FALSE)
# There is support for inclusion of Year as an RE

## Bootstrapping  to test for evidence against including multiple random effects in the same model
## set random seed 
set.seed(514)

## fit null model with no RE's 
nbm <- lm(Success ~ Type, data = SCMU)

## calculate likelihood ratio (difference in log-likelihood)
lambda <- 2 * (logLik(bm) - logLik(nbm))

## number of bootstrapped samples
nb <- 1000

## empty vector for storing LRT statistics
LRT_boot <- rep(NA, nb)

## bootstrapping
for(i in 1:nb) { # repeat nb times
  sim_data <- unlist(simulate(nbm)) # simulate data from null model
  m_null <- lm(sim_data ~ Type, data = SCMU) # fit null model to sim data 
  m_alt <- lmer(sim_data ~ Type + (1 | Year) + (1 | Plot), data = SCMU, REML = FALSE) # fit RE model
  LRT_boot[i] <- as.numeric(2*(logLik(m_alt) - logLik(m_null))) # calculate likelihood ratio 
}

## calculate approximate p-value
mean(LRT_boot > lambda)
# none of the bootstrapped samples had a test statistic larger than the original value of lambda
```

# Results


## Tables

\vspace{0.25in}

\setlength{\leftskip}{0.5in}
\setlength{\rightskip}{0.5in}

\small

`r tbl_cap(1, caption = "Model selection results.", display = "full")` 

\normalsize

\setlength{\leftskip}{0in}
\setlength{\rightskip}{0in}

```{r tbl, echo = FALSE}
tbl <- allm

# print table
knitr::kable(tbl,
             format = "latex",
             booktabs = TRUE,
             col.names = c("Model ID",
                           "Fixed Effects",
                           "$k$",
                           "$\\Delta$AIC",
                           "$\\Delta$BIC"),
             escape = FALSE,
             row.names = FALSE,
             align = "ccr") %>%
  kable_styling(position = "center", font_size = 10)
```








