---
title: "ASSP_CPUE_finalGraphs.rmd"
author: "EKelsey & ADuVall"
date: "7/8/2020"
output:
  html_document:
    df_print: paged
---
This is `r paste0("v.", (Sys.Date()))`

# STORM-PETREL MISTNETTING CPUE GRAPHS FOR FINAL NFWF REPORT
## this script creates final graphs representing CPUE across years, between sites, and related to other variables

```{r message = FALSE}
### SET WORKING DIRECTORY
# setwd("~/WERC-SC/ASSP_share")

## LOAD LIBRARIES
library(tidyr)
library(dplyr)
library(lubridate)
library(hms)
library(tidyverse)
library(ggplot2)
library(EnvStats)
library(here)
# library(calecopal)
library(cowplot)
library(anchors)
library(gridExtra)
library(pastecs)


### READ IN DATA
banding <- readRDS(here("ASSP_share", "captures.RDS"))
metadata <- readRDS(here("ASSP_share", "cpue.RDS")) 
wx_metadata <- readRDS(here("ASSP_share", "mistnet_w_weather.RDS")) 
wx_metadata_hrly <- readRDS(here("ASSP_share", "mistnet_w_weather_hrly.RDS")) 

## Filter data to ASSP species and remove SNRs and unbanded individuals 
ASSP <- banding %>% 
  filter(species == "ASSP" & recapture != "SNR" & band_no != "notbanded")  # 3799 observations
  
ASSP_BP <- ASSP %>% 
  filter(BP != "NR" & BP != "NA") %>% 
  mutate(assumeBreed = mosaic::derivedFactor(
                "Y" = (BP == "B" | BP == "b" | BP == "2" | BP == "3" | BP == "4"),
                "N" = (BP == "D" | BP == "d" | BP == "0" | BP == "5" | BP == "PD" | BP == "pd" | BP == "1" | BP == "1.5" | BP == "4.5"),
                .default = "ND")) %>% 
  left_join(metadata, by = c("site_code", "session_ID", "lat", "long")) %>% 
  mutate(std = if_else(std_ending > capture_date, "1", "0"))

metadata_BP <- ASSP_BP %>%
  group_by(session_ID, site_code) %>%
  summarise(BPct = n(),
            BPct_std = sum(std == "1"),
            BP_Y = sum(assumeBreed == "Y"), # number of birds that have a broodpatch (2-4, B)
            BP_Ystd = sum(assumeBreed == "Y" & std == "1")) %>% # number of birds that have a have a broodpatch and were caught within the std time 
  right_join(metadata, by= c("session_ID", "site_code")) %>%
  mutate(BPfreq_Y = BP_Y/BPct, # frequency of birds that have a broodpatch
         BPfreq_Ystd = BP_Ystd/BPct_std,
         CPUEbrd = BP_Y/min,
         CPUEbrd_std = BP_Ystd/min_std) 

unbanded <- banding %>%
  filter(species == "ASSP" & band_no == "notbanded") # 20 unbanded ASSP
```

# Summary Statistics
```{r summaries}
# CPUEraw
summary(metadata$CPUEraw)
stat.desc(metadata$CPUEraw)

# CPUEstd
summary(metadata$CPUEstd)
stat.desc(metadata$CPUEstd)

# broodpatch frequency
summary(metadata$BPfreq_Y)
stat.desc(metadata$BPfreq_Y)

sites <- metadata %>%
  group_by(island_code, site_name) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

months <- metadata %>%
  group_by(session_month) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

species <- banding %>%
  group_by(species) %>%
  summarize(n=n()) %>%
  arrange(desc(n))
```

#### CPUE graphs
```{r CPUE sessions}
metadata_series <- metadata %>% 
  filter(!is.na(series_ID))

series <- ggplot(metadata_series, aes(as.character(series_ID), CPUEraw)) +
  geom_boxplot() +
  stat_n_text(y.pos = 0.4) +
  xlab("Sequential Night") + ylab("Catch-Per-Unit-Effort") +
  theme_cowplot(12) # theme_bw()
series
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUEseries.pdf")

series_std <- ggplot(metadata_series, aes(as.character(series_ID), CPUEstd)) +
  geom_boxplot() +
  stat_n_text(y.pos = 0.4) +
  xlab("Sequential Night") + ylab("Standardized Catch-Per-Unit-Effort") +
  theme_cowplot(12) # theme_bw()
series_std
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUEseries_std.pdf")
```

# CPUE by Month and Year

## MONTH
```{r CPUE month 1}
month <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  stat_n_text(y.pos = 0.5) +
  xlab("Month") + ylab("Standardized Catch-Per-Unit-Effort") +
  theme_cowplot(12) # theme_bw()
month
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_month.pdf")


# gather monthly data so that raw and std CPUE can be graphed side-by-side
metadata_month <- metadata %>%
  dplyr::select(CPUEraw, CPUEstd, session_month)
metadata_month <- gather(metadata_month, "CPUEtype", "CPUE", CPUEraw:CPUEstd)

month2 <- ggplot(metadata_month, aes(as.character(session_month), CPUE)) +
  geom_boxplot() +
  facet_wrap(.~CPUEtype) +
  stat_n_text(y.pos = 0.5) +
  xlab("Month") + ylab("Catch-Per-Unit-Effort") +
  theme_cowplot(12) # theme_bw()
month2
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_month2.pdf")
```
These graphs show how catch rates varied across months 

```{r CPUE month 2}
# monthly CPUE broken out by sub-island
month_subisl <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~subisland_code) +
  xlab("Month") + ylab("Standardized CPUE") +
  stat_n_text(y.pos = 0.35) +
  theme_cowplot(12) # theme_bw()
month_subisl %+% subset(metadata, subisland_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))

# monthly CPUE broken out by island
month_isl <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~island_code) +
  xlab("Month") + ylab("Standardized CPUE") +
  stat_n_text(y.pos = 0.42) +
  theme_cowplot(12) # theme_bw()
month_isl
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_month_isl.pdf")
```
This graph shows catch rates by month for all locations combined and then each sub-island and island separately


### Month and assumed breeders
```{r CPUE month 3}
metadata_brd <- metadata_BP %>%
  dplyr::select(CPUEstd, CPUEbrd_std, session_month)
## CPUEbrd_std are brood patches w/in the std effort 

metadata_brd <- gather(metadata_brd, "CPUEtype", "CPUE", CPUEstd:CPUEbrd_std)

month_brd <- ggplot(metadata_brd, aes(as.character(session_month), CPUE, color = CPUEtype)) +
  geom_boxplot() +
  stat_n_text(y.pos = 0.4) +
  xlab("Month") + ylab("Standardized CPUE") +
  theme_cowplot(12) # theme_bw()
month_brd
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUEbrd_month.pdf")
```
This graphs shows the total CPUE each month, as well as the CPUE of assumed breeders (based on broodpatch score)
NOTE: the samples sizes at the top are the sum of CPUEbrd_std and CPUEstd

## YEAR
```{r CPUE year 1}
year <- ggplot(metadata, aes(as.character(session_year), CPUEstd)) +
  geom_boxplot() +
  xlab("Year") + ylab("Standardized Catch-Per-Unit-Effort") +
  stat_n_text(size = 3.5, y.pos = 0.41) +
  theme_cowplot(12) # theme_bw()
year %+% subset(metadata, session_year %in% c("1994", "1995", "1996", "2004", "2005", "2006", "2007", "2009", 
                                            "2010", "2011", "2014", "2015", "2016", "2017", "2018"))
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_yr.pdf")


# gather annual data by CPUE so that CPUE raw and std can be graphed alongside eachother
metadata_yr <- metadata %>%
  dplyr::select(CPUEraw, CPUEstd, session_year)
metadata_yr <- gather(metadata_yr, "CPUEtype", "CPUE", CPUEraw:CPUEstd)

year2 <- ggplot(metadata_yr, aes(as.character(session_year), CPUE)) +
  geom_boxplot() +
  facet_grid(CPUEtype~.) +
  xlab("Year") + ylab("Catch-Per-Unit-Effort") +
  stat_n_text(size = 3, y.pos = 0.37) +
  theme_bw()
year2
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_yr2.pdf")
```
This graph show how catch rates varied across years for all sites combined

```{r CPUE year 2}
# annual CPUE by subisland
year_subisl <- ggplot(metadata, aes(as.character(session_year), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~subisland_code) +
  xlab("Year") + ylab("Standardized CPUE") +
  stat_n_text(size = 2, y.pos = 0.3) +
  theme(axis.text.x = element_text(angle = 90)) 
year_subisl %+% subset(metadata, subisland_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))

# annual CPUE by island
year_isl <- ggplot(metadata, aes(as.character(session_year), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~island_code) +
  xlab("Year") + ylab("Standardized CPUE") +
  stat_n_text(size = 2, y.pos = 0.41) +
  theme(axis.text.x = element_text(angle = 90)) 
year_isl 
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_yr_isl.pdf")
```
These graphs show how catch rates vary across years for sub-island and island separately
NOTE: This graph is broken out by sub-island which is intermediate between "island" and "site".  

## year and assumed breeders
```{r CPUE year 3}
metadata_brd_yr <- metadata_BP %>%
  dplyr::select(CPUEstd, CPUEbrd_std, session_year)
metadata_brd_yr <- gather(metadata_brd_yr, "CPUEtype", "CPUE", CPUEstd:CPUEbrd_std)

year_brd <- ggplot(metadata_brd_yr, aes(as.character(session_year), CPUE, color = CPUEtype)) +
  geom_boxplot() +
  xlab("Year") + ylab("Standardized CPUE") +
  stat_n_text(size = 3, y.pos = 0.4) +
  theme(axis.text.x = element_text(angle = 90))
year_brd
ggsave(width = 10, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUEbrd_yr.pdf")
```
This graph shows annual CPUE compared to the CPUE of assumed breeders (based on broodpatch score)
NOTE: the samples sizes at the top are the sum of CPUEbrd_std and CPUEstd

## MONTH VS YEAR
```{r CPUE month year}
month_year <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~session_year) +
  xlab("Month") + ylab("Standardized CPUE") +
  stat_n_text(size = 2.5, y.pos = 0.35) +
  theme_cowplot(12) # theme_bw()
month_year %+% subset(metadata, session_year %in% c("1994", "1995", "1996", "2004", "2005", "2006", "2007", "2009", 
                                            "2010", "2011", "2014", "2015", "2016", "2017", "2018"))
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_month_yr.pdf")
```
This graph shows how catches varied across months, for each year separately.  Most years do not have enough data to show much of a pattern

## ISLANDS AND SITES 
```{r CPUE sites 1}
island <- ggplot(metadata, aes(island_code, CPUEstd)) +
  geom_boxplot() +
  xlab("Island") + ylab("Standardized Catch-Per-Unit-Effort") +
  stat_n_text(size = 4, y.pos = 0.41) +
  theme_cowplot(12) # theme_bw()
island 
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_isl.pdf")


# gather CPUE raw and std data so that island CPUE raw and std can be viewed side-by-side
metadata_isl <- metadata %>%
  dplyr::select(CPUEraw, CPUEstd, island_code, site_code)
metadata_isl <- gather(metadata_isl, "CPUEtype", "CPUE", CPUEraw:CPUEstd)

island2 <- ggplot(metadata_isl, aes(island_code, CPUE)) +
  geom_boxplot() +
  facet_wrap(.~CPUEtype) +
  xlab("Island") + ylab("Catch-Per-Unit-Effort") +
  stat_n_text(size = 4, y.pos = 0.41) +
  theme_cowplot(12) # theme_bw()
island2 
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_isl2.pdf")
```
These graphs show how catch rates varied across islands for all years combined

```{r CPUE sites 2}
site <- ggplot(metadata, aes(site_code, CPUEraw)) +
  geom_boxplot() +
  facet_wrap(.~island_code, scales = "free_x") +
  xlab("Sites Across Islands") + ylab("Raw Catch-Per-Unit-Effort") +
  stat_n_text(size = 3, y.pos = 0.4) +
  theme_cowplot(12) # theme_bw()
site  %+% subset(metadata, site_code %in% c("AP", "ESP", "PI", "SR"))
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_siteRaw.pdf")

site <- ggplot(metadata, aes(site_code, CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~island_code, scales = "free_x") +
  xlab("Sites Across Islands") + ylab("Standardized Catch-Per-Unit-Effort") +
  stat_n_text(size = 3, y.pos = 0.4) +
  theme_cowplot(12) # theme_bw()
site  %+% subset(metadata, site_code %in% c("AP", "ESP", "PI", "SR"))
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_siteStd.pdf")
```
These graphs show how catch rates varied across sites across islands for all years combined

```{r CPUE sites 2}
# select mistnetting sites that correspond with acoustic sites
metadata_ARU <- metadata %>% 
  filter(site_code == c("AP", "ESP", "SR", "PI"))

isl_yr <- ggplot(metadata_ARU, aes(as.character(session_year), CPUEstd)) +
  geom_boxplot() +
  facet_grid(.~site_code) +
  xlab("Island") + ylab("Standardized CPUE") +
  stat_n_text(size = 3, y.pos = 0.3) +
  theme_cowplot(12) # theme_bw()
isl_yr %+% subset(metadata_ARU, session_year %in% c("2016", "2017", "2018"))
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_ARUsites.pdf")
```
This graph shows how catches varied across islands in 2016 - 2018, the years in which acoustic data was collected.  This graph is comprable to the call rate outputs graphs being generated by Kerry.

### Islands and assumed breeders
```{R CPUE sites 4}
metadata_brd_isl <- metadata_BP %>%
  dplyr::select(CPUEstd, CPUEbrd_std, island_code)
metadata_brd_isl <- gather(metadata_brd_isl, "CPUEtype", "CPUE", CPUEstd:CPUEbrd_std)

island_brd <- ggplot(metadata_brd_isl, aes(island_code, CPUE, color = CPUEtype)) +
  geom_boxplot() +
  xlab("Island") + ylab("Standardized CPUEof Assumed Breeders") +
  stat_n_text(size = 4, y.pos = 0.4) +
  theme_cowplot(12) # theme_bw()
island_brd # %+% subset(metadata, site_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUEbrd_isl.pdf")
```
This graph shows the average CPUE at on each island and also the average CPUE of assumed breeders (based on broodpatch score)
NOTE: the samples sizes at the top are the sum of CPUEbrd_std and CPUEstd


## CATCH TIMES
```{r catch times 1}
# couldn't figure out how to calcuate catch time past sunset in R. Exported to excel and subtracted catch time from sunset time, then imported back:
catches_rev <- read.csv('~/WERC-SC/ASSP_share/ASSP_CPUE_7_catchesPastSS_rev.csv') %>%  
  mutate_at(c("app_sunset", "std_ending", "capture_date", "release_date"),
            .funs = ~as.POSIXct(., format="%m/%d/%Y %H:%M")) %>% # SECONDS dropped from app_sunset and std_ending during transition to and from excel
  mutate(session_date = as.POSIXct(session_date, format="%m/%d/%Y"),
         hrsPastSS = catchPastSS/60) %>% 
  filter(TRUE)

catches_test <- catches_rev %>% 
  filter(catchPastSS > 500) %>% 
  dplyr::select(catch_ID:site_code, app_sunset, capture_date, catchPastSS)
```

### catch times vs. 5.3 hours post SS, subset by year
```{r catch times 2}
endT_yr <- ggplot(catches_rev, aes(hrsPastSS)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = 5.3, color = "red") +
  # geom_vline(xintercept = 0, color = "green") +
  scale_x_continuous(name="Time past Sunset (hours)", breaks=seq(-1, 9, 1)) + 
  ylab("ASSP Catch Frequency") +
  theme_cowplot(12) # theme_bw()
endT_yr
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/catchTime_hist.pdf")
```
These graphs show the time after sunset of ASSP catches.

### catch times vs. 5.3 hours post SS, subset by island
```{r cath times 3}
endT_yr <- ggplot(catches_rev, aes(catchPastSS)) +
  geom_histogram(binwidth = 10) +
  geom_vline(xintercept = 318, color = "red") +
  geom_vline(xintercept = 0, color = "green") +
  xlab("Time past Sunset (mins)") + 
  ylab("Number of ASSP Catches") +
  facet_wrap(.~ session_year) +
  theme_bw()
endT_yr

endT_isl <- ggplot(catches_rev, aes(catchPastSS)) +
  geom_histogram(binwidth = 10) +
  geom_vline(xintercept = 318, color = "red") +
  geom_vline(xintercept = 0, color = "green") +
  xlab("Time past Sunset (min)") + ylab("Number of ASSP Catches") +
  facet_wrap(.~ subisland_code) +
  theme_bw()
endT_isl %+% subset(catches_rev, subisland_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))
```
This graph shows the time after sunset of ASSP catches, broken out by year and thens sub-island.  The green vertical line indicates sunset, the red vertical line indicates 5.3 hours after sunset, the cutoff time indicated by Adams (2016).
 

### CPUE vs. cumulative mintues
```{r CPUE total min 1}
# function to generate linear regression equation
lm_eqn <- function(df, y = y, x = x){
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}

CPUE_min <- ggplot(metadata, aes(min, CPUEraw)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  geom_text(x = 450, y = 0.25, label = lm_eqn(metadata, y = metadata$CPUEraw, x = metadata$min), parse = TRUE) + 
  xlab("Total Netting Minutes") + ylab("Catch-Per-Unit-Effort") +
  theme_cowplot(12) # 
CPUE_min
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUE_v_min.pdf")

CPUE_minstd <- ggplot(metadata, aes(min_std, CPUEstd)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  xlab("Total Netting Minutes") + ylab("Standardized Catch-Per-Unit-Effort") +
  geom_text(x = 250, y = 0.35, label = lm_eqn(metadata, y = metadata$CPUEstd, x = metadata$min_std), parse = TRUE) + 
  theme_cowplot(12) 
CPUE_minstd
ggsave(width = 7, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/CPUEstd_v_min.pdf")
```
These graphs show the CPUE (raw and std) in relation to the number of minutes the net was open.  The blue line shows the linear regression of these points.

```{r CPUE total min 2}
# broken out by year
CPUE_minstd_c <- ggplot(metadata, aes(min_std, CPUEstd, color = session_year)) +  
  geom_point() +
  xlab("Total Standardized Netting Minutes") + ylab("Standardized CPUE") +
  theme_bw()
CPUE_minstd_c
```
This graph shows the same scatter plot of standardized CPUE, broken out by year

```{r netting effort gnatt chart}
ggplot(metadata, aes(session_year, island_code, color = org)) +
  geom_point(size = 12, shape = 15) + 
  scale_x_continuous(name="Years of Mistnetting effort", breaks=seq(1994, 2018, 4)) +  
  ylab("Island") +
  scale_color_brewer(palette="BrBG") +
  theme(legend.position="bottom") #
  # theme_cowplot(12)
ggsave(width = 12, height = 5, dpi=300, filename = "~/WERC-SC/ASSP_share/figures/ganttYrs.pdf")

```
This graph shows years of netting effort broken out by organization that performed the netting

# MORPHOMETRIC DATA
### Morphometrics across islands
```{r morph}
# mass <- ggplot(data = ASSP, aes(x = island_code, y = mass_corr)) +
#   geom_boxplot() +
#   xlab("Island") + ylab("Mass (corrected)")+
#   theme_bw()
# mass

p1 <- ggplot(data = ASSP, aes(x = mass_corr), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 1, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(mass_corr, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Mass (corrected)") +
  geom_text(aes(x = mean(mass_corr, na.rm = TRUE) + 10, label= paste0("mean =\n", round(mean(mass_corr, na.rm = TRUE), 2)), y=0.1)) +
  theme_cowplot(12)

p2 <- ggplot(data = ASSP, aes(x = culmen), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(culmen, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Culmen") +
  theme_cowplot(12)

p3 <- ggplot(data = ASSP, aes(x = skull_length), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(skull_length, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Skull Length") +
  theme_cowplot(12)

p4 <- ggplot(data = ASSP, aes(x = tarsus), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(tarsus, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Tarsus") +
  theme_cowplot(12)

p5 <- ggplot(data = ASSP, aes(x = wing), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 1, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(wing, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Wing Chord") +
  theme_cowplot(12)

# plot in grid
plot_grid(p1, p2, p3, p4, p5)

# densityplot(~ mass_corr | island_code, data = ASSP)

## try boxplots
c1 <- ggplot(data = ASSP, aes(x = island_code, y = mass_corr, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Mass") +
  geom_hline(aes(yintercept = mean(mass_corr, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .65) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

c2 <- ggplot(data = ASSP, aes(x = island_code, y = culmen, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Culmen") +
  geom_hline(aes(yintercept = mean(culmen, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .65) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

# fix color palette for skull length since there are no ANI values
slc <- calecopal::cal_palette("sbchannel")[2:4]

c3 <- ggplot(data = ASSP, aes(x = island_code, y = skull_length, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Skull Length") +
  geom_hline(aes(yintercept = mean(skull_length, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) +
  scale_color_manual(values = slc) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

c4 <- ggplot(data = ASSP, aes(x = island_code, y = tarsus, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Tarsus") +
  geom_hline(aes(yintercept = mean(tarsus, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

c5 <- ggplot(data = ASSP, aes(x = island_code, y = wing, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Wing Chord") +
  geom_hline(aes(yintercept = mean(wing, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

# Look for indication of different methods used to measure wing chord 
# (e.g, flattened wing chord versus relaxed wing chord). 
# by island
wing.isl <- ggplot(ASSP, aes(x = island_code, y = wing)) +
  geom_boxplot() +
  xlab("Island") + ylab("Wing Chord (mm)") +
  geom_hline(yintercept = avwing, linetype = "dashed", color = "black", size = 1, alpha = 0.5) +
  theme_bw()

# by organization and island
ASSPorg <- inner_join(ASSP, metadata, by = "session_ID") %>%
  replace.value(., names = c("org"), from = "USGS WERC", to = "USGS") %>%
  replace.value(., names = c("org"), from = "Calif. Institute of Env'l Studies (CIES)/Channel Islands National Park (CHIS)", 
                to = "CIES/CHIS") %>%
  replace.value(., names = c("org"), from = "PRBO (Point Blue Conservation Science)", to = "PRBO") %>%
  replace.value(., names = c("org"), from = "Humboldt State University (H. Carter & D. Whitworth)", to = "HSU")

c6 <- ggplot(ASSPorg, aes(x = org, y = wing)) +
  geom_boxplot() +
  xlab("Organization") + ylab("Wing Chord") +
  geom_hline(yintercept = mean(wing, na.rm =  TRUE), linetype = "dashed", 
             color = "black", size = 1, alpha = 0.5) +
  theme_cowplot(12) +
  stat_n_text()

# plot in grid
morphs <- plot_grid(c1, c2, c3, c4, c5, c6)
# export dimensions 900x500

## table of values
morph_stats <- ASSP %>% 
  rename(skulllength = skull_length, masscorr = mass_corr) %>%
  summarise_at(vars(masscorr, culmen, skulllength, tarsus, wing), funs(mean,sd), na.rm = TRUE) %>%
  dplyr::select(order(colnames(.))) %>%
  tidyr::pivot_longer(contains("_"), # break out the measurement cols
                      names_to = c("part", "measure"), # break them into these columns
                      names_sep = "_") #  use the underscore to separate 

```

## RECAPTURES
```{r recap}
recap <- group_by(.data = ASSP, session_year, session_month, island_code) %>%
  summarise(no_captures = n(),
            recaps = sum(recapture == "Y"),
            ratio = recaps/no_captures)

# # recapture rate over time
# ggplot(recap, aes(x = as.character(session_year), y = ratio)) +
#   geom_boxplot() +
#   xlab("Year") + ylab("Recapture Rate") +
#   theme_bw()
# 
# ggplot(recap, aes(x = as.character(session_year), y = ratio, fill = island_code)) +
#   geom_col() +
#   xlab("Year") + ylab("Recapture Rate") +
#   scale_fill_manual(values = cal_palette("sierra2")) +
#   geom_hline(aes(yintercept = sum(recaps)/sum(no_captures)),
#              colour = "black", linetype ="longdash", size = .65) +
#   theme_bw()
# 
# ggplot(recap, aes(x = as.character(session_year), y = ratio)) +
#   geom_boxplot() +
#   theme_cowplot(12)

# recap by month
r1 <- ggplot(recap, aes(x = as.character(session_month), y = ratio)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Recapture Rate") +
  xlab("Month")  +
  theme_cowplot(12) +
  stat_n_text()

# recap by island
r2 <- ggplot(recap, aes(x = island_code, y = ratio)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Recapture Rate") +
  xlab("Island")  +
  theme_cowplot(12) +
  stat_n_text()

# recap by year
r3 <- ggplot(recap, aes(x = as.character(session_year), y = ratio)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Recapture Rate") +
  xlab("Year")  +
  theme_cowplot(12) +
  theme(axis.text.x = element_text(angle = 30)) +
  stat_n_text() +
  geom_hline(aes(yintercept = mean(ratio, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) 

# plot in grid
# breeders <- plot_grid(b1, b2, b3)
r <- grid.arrange(arrangeGrob(r1,r2, ncol=2, nrow=1),r3)
ggsave("recaps.jpg", r, width = 9, height = 5)

```

## BROOD PATCH
```{r broodpatch}
summary(ASSP$BP)
unique(ASSP$BP) 
summary(metadata$BPfreq_Y, na.rm = TRUE)
sd(metadata$BPfreq_Y, na.rm = TRUE)

## Brood Patch and Assumed Breeders
# breeder by month
b1 <- ggplot(metadata, aes(x = as.character(session_month), y = BPfreq_Y)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Frequency of \n Assumed Breeders") +
  # facet_wrap(.~island_code, ncol = 1) +
  xlab("Month")  +
  theme_cowplot(12) +
  stat_n_text()

# breeder by island
b2 <- ggplot(metadata, aes(x = island_code, y = BPfreq_Y)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Frequency of \n Assumed Breeders") +
  # facet_wrap(.~island_code, ncol = 1) +
  xlab("Island")  +
  theme_cowplot(12) +
  stat_n_text()

# breeder by year
b3 <- ggplot(metadata, aes(x = as.character(session_year), y = BPfreq_Y)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Frequency of \n Assumed Breeders") +
  # facet_wrap(.~island_code, ncol = 1) +
  xlab("Year")  +
  theme_cowplot(12) +
  theme(axis.text.x = element_text(angle = 30)) +
  stat_n_text() +
  geom_hline(aes(yintercept = mean(BPfreq_Y, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) 

# plot in grid
# breeders <- plot_grid(b1, b2, b3)
bp <- grid.arrange(arrangeGrob(b1,b2, ncol=2, nrow=1),b3)
ggsave("breeders.jpg", bp, width = 9, height = 5)
```