---
title: "ASSP_CPUE_finalGraphs.rmd"
author: "EKelsey & ADuVall"
date: "7/8/2020"
output: pdf_document
---
This is `r paste0("v.", (Sys.Date()))`

## STORM-PETREL MISTNETTING CPUE GRAPHS FOR FINAL NFWF REPORT
# this script creates final graphs representing CPUE across years, between sites, and related to other variables

```{r message = FALSE}
### SET WORKING DIRECTORY
# setwd("~/WERC-SC/ASSP_share")

## LOAD LIBRARIES
library(tidyr)
library(dplyr)
library(lubridate)
library(hms)
library(tidyverse)
library(ggplot2)
library(EnvStats)
library(here)
# library(calecopal)
library(cowplot)
library(anchors)
library(gridExtra)
library(pastecs)


### READ IN DATA
banding <- readRDS(here("ASSP_share", "captures.RDS"))
metadata <- readRDS(here("ASSP_share", "cpue.RDS"))

## Filter data to ASSP species and remove SNRs and unbanded individuals 
ASSP <- banding %>% 
  filter(species == "ASSP" & recapture != "SNR" & band_no != "notbanded") # 3799 observations

unbanded <- banding %>%
  filter(species == "ASSP" & band_no == "notbanded") # 20 unbanded ASSP
```

# Summary Statistics
```{r summaries}
# CPUEraw
summary(metadata$CPUEraw)
stat.desc(metadata$CPUEraw)

# CPUEstd
summary(metadata$CPUEstd)
stat.desc(metadata$CPUEstd)

# broodpatch frequency
summary(metadata$BPfreq_Y)
stat.desc(metadata$BPfreq_Y)
# summary(metadata$BPfreq_N)

sites <- metadata %>%
  group_by(island_code, site_name) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

months <- metadata %>%
  group_by(session_month) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

species <- banding %>%
  group_by(species) %>%
  summarize(n=n()) %>%
  arrange(desc(n))
```

#### CPUE graphs
# month and year

### MONTH
```{r CPUE month 1}
month <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  stat_n_text(y.pos = 0.32) +
  xlab("Month") + ylab("Standardized CPUE") +
  theme_bw()
month
```
This graph show how catch rates varied across months 

```{r CPUE month 2}
# sub-island
month_subisl <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~subisland_code) +
  xlab("Month") + ylab("Standardized CPUE") +
  stat_n_text(y.pos = 0.35) +
  theme_bw()
month_subisl %+% subset(metadata, subisland_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))

# island
month_isl <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~island_code) +
  xlab("Month") + ylab("Standardized CPUE") +
  stat_n_text(y.pos = 0.35) +
  theme_bw()
month_isl 
```
This graph shows catch rates by month for all locations combined and then each sub-island and island separately


## MONTH AND ASSUMED BREEDERS
```{r CPUE month 3}
# metadata_brd <- metadata %>% 
#   dplyr::select(CPUEstd, CPUEbrd_std, session_month)
# ## CPUEbrd_std are brood patches w/in the std effort v]
# 
# metadata_brd <- gather(metadata_brd, "CPUEtype", "CPUE", CPUEstd:CPUEbrd_std)
#   
# month_brd <- ggplot(metadata_brd, aes(session_month, CPUE, color = CPUEtype)) +
#   geom_boxplot() +
#   stat_n_text(y.pos = 0.3) +
#   xlab("Month") + ylab("Standardized CPUE") +
#   theme_bw()
# month_brd
```
This graphs shows the total CPUE each month, as well as the CPUE of assumed breeders (based on broodpatch score)
NOTE: the samples sizes at the top are the sum of CPUEbrd_std and CPUEstd

### YEAR
```{r CPUE year 1}
year <- ggplot(metadata, aes(as.character(session_year), CPUEstd)) +
  geom_boxplot() +
  xlab("Year") + ylab("Standardized CPUE") +
  stat_n_text(size = 3, y.pos = 0.35) +
  theme_bw()
year
```
This graph show how catch rates varied across years for all sites combined

```{r CPUE year 2}
# by subisland
year_subisl <- ggplot(metadata, aes(as.character(session_year), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~subisland_code) +
  xlab("Year") + ylab("Standardized CPUE") +
  stat_n_text(size = 2, y.pos = 0.3) +
  theme(axis.text.x = element_text(angle = -45)) + # not working?
  theme_bw()
year_subisl %+% subset(metadata, subisland_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))

# by island
year_isl <- ggplot(metadata, aes(as.character(session_year), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~island_code) +
  xlab("Year") + ylab("Standardized CPUE") +
  stat_n_text(size = 2, y.pos = 0.3) +
  theme(axis.text.x = element_text(angle = -45)) + # not working?
  theme_bw()
year_isl 
```
This graph shows how catch rates vary across years forsub-island separately
NOTE: This graph is broken out by sub-island which is intermediate between "island" and "site".  Using "island" or "site" for this graph might ultimately make more sense

## YEAR AND ASSUMED BREEDERS
```{r CPUE year 3}
# metadata_brd_yr <- metadata %>% 
#   select(CPUEstd, CPUEbrd_std, year)
# metadata_brd_yr <- gather(metadata_brd_yr, "CPUEtype", "CPUE", CPUEstd:CPUEbrd_std)
# 
# year_brd <- ggplot(metadata_brd_yr, aes(session_year, CPUE, color = CPUEtype)) +
#   geom_boxplot() +
#   xlab("Year") + ylab("Standardized CPUE") +
#   stat_n_text(size = 3, y.pos = 0.3) +
#   theme(axis.text.x = element_text(angle = -45))
# year_brd
```
This graph shows annual CPUE compared to the CPUE of assumed breeders (based on broodpatch score)
NOTE: the samples sizes at the top are the sum of CPUEbrd_std and CPUEstd

## MONTH VS YEAR
```{r CPUE month year}
month_year <- ggplot(metadata, aes(as.character(session_month), CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~session_year) +
  xlab("Month") + ylab("Standardized CPUE") +
  stat_n_text(size = 2, y.pos = 0.3) +
  theme_bw()
month_year
```
This graph shows how catches varied across months, for each year separately.  Most years do not have enough data to show much of a pattern

### CATCHES AT ISLANDS AND SITES THROUGH TIME
```{r CPUE sites 1}
island <- ggplot(metadata, aes(island_code, CPUEstd)) +
  geom_boxplot() +
  xlab("Island") + ylab("Standardized CPUE") +
  stat_n_text(size = 4, y.pos = 0.32) +
  theme_bw()
island 
```
This graphs show how catch rates varied across islands.
```{r CPUE sites 2}
site <- ggplot(metadata, aes(site_code, CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~island_code, scales = "free_x") +
  xlab("Sites Across Islands") + ylab("Standardized CPUE") +
  stat_n_text(size = 3, y.pos = 0.3) +
  theme_bw()
site # %+% subset(metadata, site_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))
```
This graph shows how catch rates varied across all sites within islands.
```{r CPUE sites 3}
isl_yr <- ggplot(metadata, aes(island_code, CPUEstd)) +
  geom_boxplot() +
  facet_wrap(.~session_year) +
  xlab("Island") + ylab("Standardized CPUE") +
  stat_n_text(size = 3, y.pos = 0.3) +
  theme_bw()
isl_yr %+% subset(metadata, session_year %in% c("2016", "2017", "2018"))
```
This graph shows now catches varied across islands in 2016 - 2018, the years in which acoustic data was collected.  This graph is comprable to the call rate outputs graphs being generated by Kerry.

## ISLAND AND ASSUMED BREEDERS
```{R CPUE sites 4}
# metadata_brd_isl <- metadata %>% 
#   select(CPUEstd, CPUEbrd_std, island_code)
# metadata_brd_isl <- gather(metadata_brd_isl, "CPUEtype", "CPUE", CPUEstd:CPUEbrd_std)
# 
# island_brd <- ggplot(metadata_brd_isl, aes(island_code, CPUE, color = CPUEtype)) +
#   geom_boxplot() +
#   xlab("Island") + ylab("Standardized CPUEof Assumed Breeders") +
#   stat_n_text(size = 4, y.pos = 0.3) +
#   theme_bw()
# island_brd # %+% subset(metadata, subisland_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))
```
This graph shows the average CPUE at on each island and also the average CPUE of assumed breeders (based on broodpatch score)
NOTE: the samples sizes at the top are the sum of CPUEbrd_std and CPUEstd


#### CATCH TIMES
```{r catch times 1}
metadata_effort <- metadata %>% 
  select(site_code, session_ID, app_sunset, std_ending, lat, long)

catches <-  banding %>% 
  left_join(metadata_effort, by = c("site_code", "session_ID", "lat", "long")) %>%
  mutate(catchPastSS = app_sunset - capture_date)
   
  # std = if_else(std_ending > capture_date, "1", "0"),
  # #        captureT = hms(capture_date),
```

## catch times vs. 5.3 hours post SS, subset by year
```{r catch times 2}
endT_yr <- ggplot(catches, aes(catchPastSS)) +
  geom_histogram(binwidth = 10) +
  geom_vline(xintercept = 318, color = "red") +
  geom_vline(xintercept = 0, color = "green") +
  xlab("Time past Sunset (min)") + ylab("Number of ASSP Catches") +
  facet_wrap(.~ session_year) +
  theme_bw()
endT_yr
```
This graph shows the time after sunset of ASSP catches, broken out by year.  The green vertical line indicates sunset, the red vertical line indicates 5.3 hours after sunset, the cutoff time indicated by Adams (2016).
NOTE: in some years there were catches recorded before sunset.  This probably isn't accurate and is a bug in how time was calcualted in R.  Needs review. 

## catch times vs. 5.3 hours post SS, subset by island
```{r cath times 3}
endT_isl <- ggplot(catches, aes(catchPastSS)) +
  geom_histogram(binwidth = 10) +
  geom_vline(xintercept = 318, color = "red") +
  geom_vline(xintercept = 0, color = "green") +
  xlab("Time past Sunset (min)") + ylab("Number of ASSP Catches") +
  facet_wrap(.~ subisland_code) +
  theme_bw()
endT_isl %+% subset(catches, subisland_code %in% c("DR", "EAI", "PI", "SBI", "SR", "Sutil"))
```
This graph shows the time after sunset of ASSP catches, broken out by sub-island.  The green vertical line indicates sunset, the red vertical line indicates 5.3 hours after sunset, the cutoff time indicated by Adams (2016).
NOTE: in some years there were catches recorded before sunset.  This probably isn't accurate and is a bug in how time was calcualted in R.  Needs review. 

## CPUE vs. cumulative mintues
```{r CPUE total min 1}
CPUE_min <- ggplot(metadata, aes(min, CPUEraw)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  xlab("Total Netting Minutes") + ylab("CPUE") +
  theme_bw()
CPUE_min
```
This graphs show the CPUE in relation to the number of minutes the net was open.  The blue line shows the linear regression of these points.
```{r CPUE total min 2}
CPUE_minstd <- ggplot(metadata, aes(min_std, CPUEstd)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  xlab("Total Netting Minutes") + ylab("Standardized CPUE") +
  theme_bw()
CPUE_minstd
```
this graph shows the standardized CPUE (until 5.3 hours after sunset) in relation to the total number of standardized minutes the net was open. The blue line shows the linear regression of these points. 
```{r CPUE total min 3}
CPUE_minstd_c <- ggplot(metadata, aes(min_std, CPUEstd, color = session_year)) + # 
  geom_point() +
  # geom_smooth(method = 'lm') +
  xlab("Total Standardized Netting Minutes") + ylab("Standardized CPUE") +
  # scale_fill_brewer(palette="Dark2")
  theme_bw()
CPUE_minstd_c
```
This graph shows the same scatter plot of standardized CPUE, broken out by color

## MORPHOMETRIC DATA
### Morphometrics across islands
```{r morph}
# mass <- ggplot(data = ASSP, aes(x = island_code, y = mass_corr)) +
#   geom_boxplot() +
#   xlab("Island") + ylab("Mass (corrected)")+
#   theme_bw()
# mass

p1 <- ggplot(data = ASSP, aes(x = mass_corr), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 1, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(mass_corr, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Mass (corrected)") +
  geom_text(aes(x = mean(mass_corr, na.rm = TRUE) + 10, label= paste0("mean =\n", round(mean(mass_corr, na.rm = TRUE), 2)), y=0.1)) +
  theme_cowplot(12)

p2 <- ggplot(data = ASSP, aes(x = culmen), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(culmen, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Culmen") +
  theme_cowplot(12)

p3 <- ggplot(data = ASSP, aes(x = skull_length), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(skull_length, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Skull Length") +
  theme_cowplot(12)

p4 <- ggplot(data = ASSP, aes(x = tarsus), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 0.5, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(tarsus, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Tarsus") +
  theme_cowplot(12)

p5 <- ggplot(data = ASSP, aes(x = wing), na.rm = TRUE) +
  geom_histogram(aes(y = ..density..), binwidth = 1, color = "white") +
  geom_density(alpha = 0.5, fill = "cadetblue") +
  geom_vline(aes(xintercept = mean(wing, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .8) +
  xlab("Wing Chord") +
  theme_cowplot(12)

# plot in grid
plot_grid(p1, p2, p3, p4, p5)

# densityplot(~ mass_corr | island_code, data = ASSP)

## try boxplots
c1 <- ggplot(data = ASSP, aes(x = island_code, y = mass_corr, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Mass") +
  geom_hline(aes(yintercept = mean(mass_corr, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .65) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

c2 <- ggplot(data = ASSP, aes(x = island_code, y = culmen, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Culmen") +
  geom_hline(aes(yintercept = mean(culmen, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .65) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

# fix color palette for skull length since there are no ANI values
slc <- calecopal::cal_palette("sbchannel")[2:4]

c3 <- ggplot(data = ASSP, aes(x = island_code, y = skull_length, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Skull Length") +
  geom_hline(aes(yintercept = mean(skull_length, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) +
  scale_color_manual(values = slc) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

c4 <- ggplot(data = ASSP, aes(x = island_code, y = tarsus, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Tarsus") +
  geom_hline(aes(yintercept = mean(tarsus, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

c5 <- ggplot(data = ASSP, aes(x = island_code, y = wing, color = island_code)) +
  geom_boxplot() +
  xlab("Island") + ylab("Wing Chord") +
  geom_hline(aes(yintercept = mean(wing, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) +
  scale_color_manual(values = cal_palette("sbchannel")) +
  theme_cowplot(12) +
  theme(legend.position = "none") +
  stat_n_text()

# Look for indication of different methods used to measure wing chord 
# (e.g, flattened wing chord versus relaxed wing chord). 
# by island
wing.isl <- ggplot(ASSP, aes(x = island_code, y = wing)) +
  geom_boxplot() +
  xlab("Island") + ylab("Wing Chord (mm)") +
  geom_hline(yintercept = avwing, linetype = "dashed", color = "black", size = 1, alpha = 0.5) +
  theme_bw()

# by organization and island
ASSPorg <- inner_join(ASSP, metadata, by = "session_ID") %>%
  replace.value(., names = c("org"), from = "USGS WERC", to = "USGS") %>%
  replace.value(., names = c("org"), from = "Calif. Institute of Env'l Studies (CIES)/Channel Islands National Park (CHIS)", 
                to = "CIES/CHIS") %>%
  replace.value(., names = c("org"), from = "PRBO (Point Blue Conservation Science)", to = "PRBO") %>%
  replace.value(., names = c("org"), from = "Humboldt State University (H. Carter & D. Whitworth)", to = "HSU")

c6 <- ggplot(ASSPorg, aes(x = org, y = wing)) +
  geom_boxplot() +
  xlab("Organization") + ylab("Wing Chord") +
  geom_hline(yintercept = mean(wing, na.rm =  TRUE), linetype = "dashed", 
             color = "black", size = 1, alpha = 0.5) +
  theme_cowplot(12) +
  stat_n_text()

# plot in grid
morphs <- plot_grid(c1, c2, c3, c4, c5, c6)
# export dimensions 900x500

## table of values
morph_stats <- ASSP %>% 
  rename(skulllength = skull_length, masscorr = mass_corr) %>%
  summarise_at(vars(masscorr, culmen, skulllength, tarsus, wing), funs(mean,sd), na.rm = TRUE) %>%
  dplyr::select(order(colnames(.))) %>%
  tidyr::pivot_longer(contains("_"), # break out the measurement cols
                      names_to = c("part", "measure"), # break them into these columns
                      names_sep = "_") #  use the underscore to separate 

```

## RECAPTURES
```{r recap}
recap <- group_by(.data = ASSP, session_year, session_month, island_code) %>%
  summarise(no_captures = n(),
            recaps = sum(recapture == "Y"),
            ratio = recaps/no_captures)

# # recapture rate over time
# ggplot(recap, aes(x = as.character(session_year), y = ratio)) +
#   geom_boxplot() +
#   xlab("Year") + ylab("Recapture Rate") +
#   theme_bw()
# 
# ggplot(recap, aes(x = as.character(session_year), y = ratio, fill = island_code)) +
#   geom_col() +
#   xlab("Year") + ylab("Recapture Rate") +
#   scale_fill_manual(values = cal_palette("sierra2")) +
#   geom_hline(aes(yintercept = sum(recaps)/sum(no_captures)),
#              colour = "black", linetype ="longdash", size = .65) +
#   theme_bw()
# 
# ggplot(recap, aes(x = as.character(session_year), y = ratio)) +
#   geom_boxplot() +
#   theme_cowplot(12)

# recap by month
r1 <- ggplot(recap, aes(x = as.character(session_month), y = ratio)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Recapture Rate") +
  xlab("Month")  +
  theme_cowplot(12) +
  stat_n_text()

# recap by island
r2 <- ggplot(recap, aes(x = island_code, y = ratio)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Recapture Rate") +
  xlab("Island")  +
  theme_cowplot(12) +
  stat_n_text()

# recap by year
r3 <- ggplot(recap, aes(x = as.character(session_year), y = ratio)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Recapture Rate") +
  xlab("Year")  +
  theme_cowplot(12) +
  theme(axis.text.x = element_text(angle = 30)) +
  stat_n_text() +
  geom_hline(aes(yintercept = mean(ratio, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) 

# plot in grid
# breeders <- plot_grid(b1, b2, b3)
r <- grid.arrange(arrangeGrob(r1,r2, ncol=2, nrow=1),r3)
ggsave("recaps.jpg", r, width = 9, height = 5)

```

## BROOD PATCH
```{r broodpatch}
summary(ASSP$BP)
unique(ASSP$BP) 
summary(metadata$BPfreq_Y, na.rm = TRUE)
sd(metadata$BPfreq_Y, na.rm = TRUE)

## Brood Patch and Assumed Breeders
# breeder by month
b1 <- ggplot(metadata, aes(x = as.character(session_month), y = BPfreq_Y)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Frequency of \n Assumed Breeders") +
  # facet_wrap(.~island_code, ncol = 1) +
  xlab("Month")  +
  theme_cowplot(12) +
  stat_n_text()

# breeder by island
b2 <- ggplot(metadata, aes(x = island_code, y = BPfreq_Y)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Frequency of \n Assumed Breeders") +
  # facet_wrap(.~island_code, ncol = 1) +
  xlab("Island")  +
  theme_cowplot(12) +
  stat_n_text()

# breeder by year
b3 <- ggplot(metadata, aes(x = as.character(session_year), y = BPfreq_Y)) +
  geom_boxplot(na.rm = TRUE) +
  ylab("Frequency of \n Assumed Breeders") +
  # facet_wrap(.~island_code, ncol = 1) +
  xlab("Year")  +
  theme_cowplot(12) +
  theme(axis.text.x = element_text(angle = 30)) +
  stat_n_text() +
  geom_hline(aes(yintercept = mean(BPfreq_Y, na.rm = TRUE)),
             colour = "black", linetype ="longdash", size = .5) 

# plot in grid
# breeders <- plot_grid(b1, b2, b3)
bp <- grid.arrange(arrangeGrob(b1,b2, ncol=2, nrow=1),b3)
ggsave("breeders.jpg", bp, width = 9, height = 5)
```