---
title: "CPUE"
author: "by Julie Yee, for Emma Kelsey"
date: "December 28, 2016"
output: html_document
---

##Background

**Aug 15, 2016 email from Emma Kelsey**
I have a statistical question- I want to do a regression analysis between 5 variables where the dependent variable is categorical (5 categories).  I am working in R.  

**Aug 16, 2016 email from Emma Kelsey**
I've been looking into multinomial logistic regression models and I have a couple questions:
First of all, most of the information I've been finding about multinomial logistic regression models is to use them as to generate predictions.  I am just interested in understanding which of my four independent variables best explains the variation in my dependent variable.  Whats the difference?  Or is there one?
This example (<http://www.ats.ucla.edu/stat/r/dae/mlogit.htm>) is the one that makes the most sense to me, but I'm not sure of it applies to what I'm trying to do.
Here's a little more about my data set:
I am looking at 14 years of invasive mammal control data at Haleakala National Park, Maui (2000-2014).  There are 18 trap lines set up within the park with traps to catch invasive predators (cats, rats, mongoose, and mice).  The trap lines have a number of traps alon g them that are checked each week.  When the traps are checked the possible trap status is: cat caught, rat caught, mongoose caught, mouse caught, bait lost (trap empty & open), trap triggered ( trap empty & closed), or none (trap open, bait still there).  I am interested in which variables: month of the year, year (2000-2014), bait type, or trap line best explains the trap status.  

**Sep 26, 2016 email from Emma Kelsey**
The dataset that I am working has to do with the trapping of invasive predators within a national park.  We are trying to assess the "catch per unit effort" (CPUE, effort = weekly trap checks) of different predators (cat, rat, mongoose, mouse) along 19 transects.  Data was collected 2000 - 2014.
I am attaching the datasheet here. 
What I am trying to do is run a repeated-measures ANOVA for the relationship between CPUE (the frequency of "predEvent" per "NTraps" for each week) and Trapline, Year, Season. The data is not normally distributed, so I think I need to run a Friedman's test.  I am not sure how to do this and would love some help.  Please feel free to call me: 253-219-7165

**Oct 21, 2016 email from Emma Kelsey**
I am (finally) getting back to working on this analysis.  Max helped me add the zeros back into the data and now I am trying to determine the best model to use, based  on your advice.
I have a couple questions following up on what we discussed:
What is the best way to add random effects into the glm?  I agree that Trapline (and perhaps Year?) should be random effects.  If I make them random effects, can I still assess their effect on predEvent per unit effort?
In trying to stay focused on the question I'm trying to ask - ***does the predEvent per unit effort vary between different traplines, years and seasons?***
I'm realizing I need to incorporate PredEvent into the model in some way, because each type of predEvent might have a different relationship with trapline, year, or season.  At first I was thinking that I could run the model on each predEvent separately, but this would be biased (e.g. - If the trap is triggered, then that excludes to possibility of a cat/rat/mogoose/mouse caught).  Could you help me figure out how to incorporate predEvent into the model this way? 

###Objective
The question is ***Does the predEvent per unit effort vary between different traplines, years and seasons?***  For these data, there is only one predEvent outcome per trap, since a trap that captures one individual will not be able to capture another.  Another way to ask the same question is ***Does the probability of predEvent vary between different traplines, years and seasons?***

## What is best model for trapline analysis?
I'd go with the `mlogit` package for its ability to handle random effects, which seem problematic for the Poisson log-linear modeling approach and perhaps impossible for the `multinom` function.  

## Emma Kelsey's catch-per-unit-effort data
```{r}
library(XLConnect)
data <- readWorksheetFromFile("TraplinePredEventPUE_11.xlsx",
                              sheet="TraplinePredEventPUE_11")
summary(data)
dim(data)
with(data, table(predEvent))
with(data, table(Trapline, predEvent))
with(data, table(Year, predEvent))
with(data, table(Season, predEvent))
with(data, table(Trapline, predEvent, Season))
```
Number of traps range from 1 to 81.  For every trap that is set, there are 8 possible outcomes (baitLost, birdOtherCaught, catCaught, mongooseCaught, mouseCaught, ratCaught, trapTriggered, none).  There are 19 different traplines, 15 different years, and 4 different seasons.  

***Data needs to be reshaped before proceeding...***
According to vignette, "Data sets can have two different shapes : 

+ a wide shape : in this case, there is one row for each choice situation,
+ a long shape : in this case, there is one row for each alternative and, therefore, as many rows as there are alternatives for each choice situation."

Apparently, we need to reshape the data so that there is one row for every option, for every choice situation.  Here are the possible outcomes every time the trap is set:
```{r}
(predEvents <- unique(data$predEvent))
```

And here are the number of choice situations:
```{r}
(nEvents <- sum(data$NEvents))
```

There are `r length(predEvents)` possible outcomes every time trap was set, and trap was set `r nEvents` times.  Total rows in data would be the product `r nEvents*length(predEvents)`.  

Perform in two steps.  First, expand the rows so that each choice situation is on its own unique row.  
```{r}
library(plyr)
library(dplyr)
```

```{r}
data2 <- data[rep(row.names(data), data$NEvents),]
data2 <- data2 %>%
  mutate(chid = row.names(data2))
```

Then expand each choice situation so that each alternative is on its own row.  Do this with the merge function.  The alternative names will be stored in column `x`.
```{r}
expanded_data <- merge(predEvents, data2)
expanded_data <- expanded_data %>% 
    mutate(choice = ifelse(x==predEvent, TRUE, FALSE)) 
```

Now apply the `mlogit.data` function.  
```{r}
library(mlogit)
cpue2 <- mlogit.data(expanded_data, choice="choice",
                     alt.var ="x", 
                     shape="long", chid.var="chid")
```

Simple `mlogit` model where Trapline and Season are specific to the choice situation, or "individual"-specific.  I.e. not alternative-specific.      
```{r, eval=FALSE}
cpue.models <- list()
cpue.models[[1]] <- mlogit(choice ~ 0 | Trapline + Season , data=cpue2)
```

Possible problem zero-valued cells?  See what happens when data are restricted to traplines where every outcome occurred at least once.  
```{r, eval=FALSE}
cpue3 <- mlogit.data(expanded_data %>% 
                       filter(Trapline %in% c('A','B','C','D','E','F')), 
                     choice="choice",
                     alt.var ="x", 
                     shape="long", chid.var="chid")
cpue.models[[2]] <- mlogit(choice ~ 0 | Season, data=cpue3)
```

```{r}
cpue.models[[4]] <- mlogit(choice ~ 0 | Season + Trapline + Year,
                           rpar=c(Year='n', Trapline='n'),
                           data=cpue3)
```

####December 28 email
```{r}
newdata <- read.csv("TraplinePredEventPUE_11_20161209.csv")
summary(newdata)
summary(expanded_data)

cpue4 <- mlogit.data(expanded_data2 %>% 
                       filter(Trapline %in% c('A','B','C','D','E','F')), 
                     choice="choice",
                     alt.var ="x", 
                     shape="long", chid.var="chid")
cpue.models[[4]] <- mlogit(choice ~ 0 | Season + Trapline + Year,
                           rpar=c(Year='n', Trapline='n'), halton=NA,
                           data=cpue3)

```

I searched online using this search phrase: "mlogit Error in names(sup.coef) <- names.sup.coef : 'names' attribute [1] must be the same length as the vector [0]" and this page popped up with a helpful explanation and a potential work-around involving another package:
<http://stackoverflow.com/questions/22020050/multinomial-mixed-logit-model-mlogit-r-package>

####January 19
Based on reading of 2 possible solutions (http://stats.stackexchange.com/questions/51148/unable-to-provide-random-parameter-with-mlogit) and a link to some helpful-looking training materials for the mlogit package (http://facweb.knowlton.ohio-state.edu/pviton/courses2/crp5700/5700-mlogit.pdf).  

```{r}
cpue4 <- mlogit.data(expanded_data2 %>% 
                       filter(Trapline %in% c('A','B','C','D','E','F')), 
                     choice="choice",
                     alt.var ="x", 
                     shape="long", chid.var="chid")
cpue.models[[5]] <- mlogit(choice ~ Season + Trapline + Year,
                           rpar=c(Year='n', Trapline='n'), halton=NA,
                           data=cpue4)
```
Error in solve.default(H, g[!fixed]) : 
  system is computationally singular: reciprocal condition number = 1.17961e-33

####February 15.
I just read the Viton tutorial.  I will try again starting with a model without random effects...
```{r}
cpue.models[[6]] <- mlogit(choice ~ Season + Trapline + Year,
                           data=cpue4)
```
Error in solve.default(H, g[!fixed]) : 
  system is computationally singular: reciprocal condition number = 1.17961e-33

```{r}
cpue.models[[6]] <- mlogit(choice ~ 0 | Season + Trapline + Year,
                           data=cpue4)
summary(cpue.models[[6]])
```
This worked, but is not really what Emma & I are looking for, otherwise the next logical step would be to specify coefficients to be normal, e.g.
```{r}
cpue.models[[7]] <- mlogit(choice ~ 0 | Season + Trapline + Year,
                           rpar=c('birdOtherCaught:SeasonNestling'='n'), halton=NA,
                           data=cpue4)
summary(cpue.models[[7]])
```
which takes a looong time (>30 minutes) and would fit a different normal distribution to each.  We want year and/or traplines to be random effects, but instead it is trying to generate a fixed parameter for each.  To do this, we need to assign id.var to year, trapline, or combination of both.  This would be done in the mlogit.data step.


```{r}
cpue.year <- 
  mlogit.data(expanded_data2 %>% 
                filter(Trapline %in% c('A','B','C','D','E','F')), 
              choice="choice",
              alt.var ="x", 
              id.var = "Year",
              shape="long", 
              chid.var="chid")

cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         rpar=c('birdOtherCaught:(intercept)'='n'), R=50, halton=NA,
         panel=TRUE,
         data=cpue.year)
```
I get this error message:
Error in if (abs(x - oldx) < ftol) { : 
  missing value where TRUE/FALSE needed
  
Viton 2015 in his Appendix D says this type of error occurs with very large panel data, due to small probabilities being raised to large powers.  To verify, he recommends using the `iterlim` and `print.level` options to see whether the initial value is positive $\infty$. 

```{r}
cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         rpar=c('birdOtherCaught:(intercept)'='n'), R=50, halton=NA,
         panel=TRUE,
         iterlim=1, print.level=1,
         data=cpue.year)
```
Apparently it is, and the only solution appears to be to take a subset.  Perhaps we can try subsetting by year....
   
```{r}
cpue.TrapA <- 
  mlogit.data(expanded_data2 %>% 
                filter(Trapline %in% c('A')), 
              choice="choice",
              alt.var ="x", 
              id.var = "Year",
              shape="long", 
              chid.var="chid")

cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         #rpar=c('birdOtherCaught:(intercept)'='n'), R=50, halton=NA,
         #panel=TRUE,
         #iterlim=1, print.level=1,
         data=cpue.TrapA)
summary(cpue.models[[8]])

cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         rpar=c('birdOtherCaught:(intercept)'='n'), R=50, halton=NA,
         panel=TRUE,
         iterlim=1, print.level=1,
         data=cpue.TrapA)
```
The problem persists. 




### Try another approach
Viton 2015 suggests subset sampling.  I think we can reduce the data a different way, by restrict the data to "Caught" outcomes, which are relatively low probability.  

```{r}
data.Caughts_only <- data %>% 
  filter(predEvent %in% c("catCaught","mongooseCaught","ratCaught","birdOtherCaught"))
```

Repeat the steps from above, but create a function to do this.  
```{r}
formatData <- function(data){
  # Reshape data so that there is one row for every option, for every choice situation.  
  # Here are the possible outcomes every time the trap is set:
  predEvents <- unique(data.Caughts_only$predEvent)

  # And here are the number of choice situations:
  nEvents <- sum(data$NEvents)
  
  # Replicate the rows according to number of events:
  data2 <- data[rep(row.names(data), data$NEvents),]
  data2 <- data2 %>%
    mutate(chid = row.names(data2))

  # Expand each choice situation so that each alternative is on its own row.  
  # Do this with the merge function.  The alternative names will be stored in column `x`.
  expanded_data <- merge(predEvents, data2)
  expanded_data <- expanded_data %>% 
    mutate(choice = ifelse(x==predEvent, TRUE, FALSE)) 
  return(expanded_data)
}

expanded_data.Caughts_only <- formatData(data.Caughts_only)
with(expanded_data.Caughts_only %>% filter(choice==TRUE),
     table(predEvent, Year))
with(expanded_data.Caughts_only %>% filter(choice==TRUE & Trapline=='A'),
     table(predEvent, Year))
```

Now try modeling it.  
```{r}
cpue.TrapA <- 
  mlogit.data(expanded_data.Caughts_only %>% 
                filter(Trapline %in% c('A')), 
              choice="choice",
              alt.var ="x", 
              id.var = "Year",
              shape="long", 
              chid.var="chid")

with(expanded_data.Caughts_only, table(predEvent))

cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         #rpar=c('birdOtherCaught:(intercept)'='n'), R=50, halton=NA,
         #panel=TRUE,
         #iterlim=1, print.level=1,
         data=cpue.TrapA)
summary(cpue.models[[8]])

cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         rpar=c('catCaught:(intercept)'='n'), R=50, halton=NA,
         panel=TRUE,
         #iterlim=1, print.level=1,
         data=cpue.TrapA)
```
Too few traplines?  Try again with all traplines

```{r}
cpue.Year <- 
  mlogit.data(expanded_data.Caughts_only, 
              choice="choice",
              alt.var ="x", 
              id.var = "Year",
              shape="long", 
              chid.var="chid")

cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         #rpar=c('birdOtherCaught:(intercept)'='n'), R=50, halton=NA,
         #panel=TRUE,
         #iterlim=1, print.level=1,
         data=cpue.Year)
summary(cpue.models[[8]])

cpue.models[[8]] <- 
  mlogit(choice ~ 0 | Season,
         rpar=c('catCaught:(intercept)'='n',
                'mongooseCaught:(intercept)'='n'), R=50, halton=NA,
         panel=TRUE,
         #iterlim=1, print.level=1,
         data=cpue.Year)
summary(cpue.models[[8]])
AIC(cpue.models[[8]])
```

```{r}
cpue.trap <- 
  mlogit.data(expanded_data2 %>% 
                filter(Trapline %in% c('A','B','C','D','E','F')), 
              choice="choice",
              alt.var ="x", 
              id.var = "Trapline",
              shape="long", 
              chid.var="chid")
```

```{r}
cpue.trapyr <- 
  mlogit.data(expanded_data2 %>% 
                filter(Trapline %in% c('A','B','C','D','E','F')) %>%
                mutate(trapyr=paste0(Trapline,'-',Year)),
              choice="choice",
              alt.var ="x", 
              id.var = "trapyr",
              shape="long", 
              chid.var="chid")
```
