---
title: "PFSH Distribution Model + UD"
author: "Max Czapanskiy"
date: "September 14, 2016"
output: html_document
---

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 120)
```

## Alternative Example
As with SOSH, PFSH. Smaller samples, so only model months 9 and 10 in a combined model with month as a random effect.

## Method
### Model without UD

1. Run global geographic and oceanographic models.
2. Cull models.
3. Combine culled models.

```{r load_pkgs_data, message = FALSE, warning = FALSE}
library(gamlss)
library(dplyr)
library(foreach)
library(iterators)
library(knitr)

pfsh.data <- read.csv('TelemetryTransect17May2016_SOSH-PFSH-COMU.csv') %>% 
  dplyr::select(PFSHcount, Binarea, Month, 
                Latitude, DistCoast, Dist200, DepCI,
                SSTmean, STD_SST, MEAN_Beaufort, L10CHLproxy, STD_CHL_log10, Watermass,
                L10CHLsurvey, CHLsurvanom, L10CHLsurvclim, FCPI,
                PFSH0610UD, PFSH0910UD = pfsh0910MN) %>%
  filter(Month > 7) %>% # Only 9, 10
  mutate(PFSHcount = as.integer(PFSHcount),
         Month = factor(Month)) %>% 
  na.omit
```

### Run global and oceanographic
```{r base_models, message = FALSE, warning = FALSE, cache = TRUE}
geo.global <- gamlss(PFSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)) + random(Month), 
                     family = SI, 
                     data = pfsh.data,  
                     control = gamlss.control(n.cyc = 100))
geo.global

ocean.global <- gamlss(PFSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                         cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                         cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)) + random(Month), 
                       family = SI, 
                       data = pfsh.data,  
                       control = gamlss.control(n.cyc = 100))
ocean.global
```

### Cull models
Variable culling process, as before
```{r cull_models, message = FALSE, warning = FALSE, cache = TRUE}
# Recursively drop parameters by largest decrease in AIC until model degrades
cull.model <- function(model) {
  aic <- extractAIC(model)[2]
    
  # Re-fit model to n-1 terms
  model.terms <- model %>% formula %>% terms
  term.labels <- attr(model.terms, 'term.labels')
  submodels <- foreach(i = seq(term.labels)) %do% {
    dropped <- term.labels[i]
    submodel <- try(update(model, reformulate(sprintf('. - %s', dropped))))
    list(dropped = dropped,
         submodel = submodel)
  }
  
  # Which submodels converged? Which submodel has the greatest decrease in AIC?
  drop.results <- foreach(i = seq(submodels), .combine = rbind) %do% {
    tryCatch(with(submodels[[i]], data.frame(dropped = dropped, 
                                             i = i,
                                             converged = submodel$converged,
                                             AIC = extractAIC(submodel)[2],
                                             deltaAIC = extractAIC(submodel)[2] - aic)),
             error = function(e) data.frame(dropped = NA,
                                            i = NA,
                                            converged = NA,
                                            AIC = NA,
                                            deltaAIC = NA))
  } %>%
    filter(converged) %>%
    arrange(deltaAIC)
  
  # If no submodels converge or if no submodels are an improvement, return the original model
  if(nrow(drop.results) == 0 || min(drop.results$deltaAIC, na.rm = TRUE) > 0) {
    return(model)
  } else {
    # Otherwise, repeat the process on the best submodel
    best.submodel <- submodels[[drop.results$i[1]]]$submodel
    return(cull.model(best.submodel))
  }
}

geo.culled <- cull.model(geo.global)
geo.culled

ocean.culled <- cull.model(ocean.global)
ocean.culled
```

### Combine
```{r random_effects, message = FALSE, warning = FALSE, cache = TRUE}
geo.formula <- geo.culled %>% terms %>% attr('term.labels')
ocean.formula <- ocean.culled %>% terms %>% attr('term.labels')
combined.rs <- paste(c(geo.formula, ocean.formula), collapse = ' + ')
combined.formula <- as.formula(paste("PFSHcount ~", combined.rs))
combined.culled <- gamlss(combined.formula,
                          family = SI, 
                       data = pfsh.data,  
                       control = gamlss.control(n.cyc = 100))
combined.culled
```

## Summarize models
```{r base_summaries, cache = TRUE}
formula.to.char <- function(model) {
  model %>% formula %>% terms.formula %>% attr('term.labels') %>% paste(collapse = ' + ')
}

# Utility function for calculating AIC weight. vAIC is a vector of AIC values
calcAICw <- function(vAIC) {
  deltaAIC <- vAIC - min(vAIC, na.rm = TRUE)
  relLikelihood <- exp(-0.5 * deltaAIC)
  normalizingFactor <- sum(relLikelihood, na.rm = TRUE)
  relLikelihood / normalizingFactor
}

# For all combinations of month, type, and scope...
base.summary.0 <- expand.grid(type = c('geo', 'ocean', 'combined'),
                               scope = c('global', 'culled'),
                               stringsAsFactors = FALSE) %>%
  filter(!(type == 'combined' & scope == 'global')) %>% # There are no combined global models 
  mutate(Model = paste(type, scope)) %>%
  arrange(type, scope)

base.summary <- foreach(row = iter(base.summary.0, by = 'row'), .combine = rbind) %do% {
  with(row, {
    # Fetch model
    model <- try(get(sprintf('%s.%s', type, scope)))
    
    # Verify model exists
    if(class(model)[1] == 'try-error' ||
       is.null(model))
      return(mutate(row,
                    Formula = NA,
                    gAIC = NA,
                    EDF = NA))

    # Run model diagnostics
    mutate(row,
           Formula = formula.to.char(model),
           gAIC = extractAIC(model)[2],
           EDF = extractAIC(model)[1])
  })
} %>%
  # Run model group diagnostics
  mutate(deltaGAIC = gAIC - min(gAIC, na.rm = TRUE),
         GAICw = calcAICw(gAIC)) %>%
  select(Model, Formula, gAIC, EDF, deltaGAIC, GAICw) %>%
  arrange(gAIC)

kable(base.summary, digits = 2)
```

## Best Model Diagnostics
```{r best_diagnostics, cache = TRUE}
# "type scope" -> "type.scope"
model.name <- base.summary$Model[1]
best.model <- get(sub(' ', '.', model.name))

# Coefficients
summary(best.model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(best.model)
qqnorm(residx, 
       main = "Normal Q-Q Plot",
       sub = model.name,
       xlab = "Theoretical Quantiles",
       ylab = "Sample Quantiles", 
       plot.it = TRUE, 
       frame.plot = TRUE, 
       col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(best.model)

# Term Plots
term.plot(best.model)
```

## Augment best model with UD variable

1. Season specific (9 and 10) and full survey period (6 – 10) shearwater BBUDs added to the top-ranking model to evaluate potential for increasing model fit based on gAIC.
2. What works best?

```{r add_ud, message = FALSE, cache = TRUE}
ud.seasonal <- update(best.model, ~ . + PFSH0910UD)
ud.seasonal

ud.total <- update(best.model, ~ . + PFSH0610UD)
ud.total
```

```{r ud_summaries, cache = TRUE}
ud.summary <- foreach(model = list(best.model, ud.seasonal, ud.total),
                      model.name = c(model.name, 'UD Seasonal', 'UD Total'),
                      .combine = rbind) %do% {
  # Run model diagnostics
  data.frame(Model = model.name,
             Formula = formula.to.char(model),
             gAIC = extractAIC(model)[2],
             EDF = extractAIC(model)[1],
             Converged = model$converged)
} %>%
  # Run model group diagnostics
  mutate(deltaGAIC = gAIC - min(gAIC, na.rm = TRUE),
         GAICw = calcAICw(gAIC)) %>%
  select(Model, Formula, gAIC, EDF, deltaGAIC, GAICw, Converged) %>%
  arrange(gAIC)

kable(ud.summary, digits = 2)
```

## UD Model Diagnostics
### UD Seasonal
```{r udseasonal_diagnostics, cache = TRUE}
model <- ud.seasonal

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, 
       main = "Normal Q-Q Plot",
       sub = "UD Seasonal",
       xlab = "Theoretical Quantiles",
       ylab = "Sample Quantiles", 
       plot.it = TRUE, 
       frame.plot = TRUE, 
       col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```

### UD Total
```{r udtotal_diagnostics}
model <- ud.total

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, 
       main = "Normal Q-Q Plot",
       sub = "UD Total",
       xlab = "Theoretical Quantiles",
       ylab = "Sample Quantiles", 
       plot.it = TRUE, 
       frame.plot = TRUE, 
       col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```
