---
title: "SOSH GAMLSS"
author: "Max Czapanskiy"
date: "July 12, 2016"
output: html_document
---

Goal: demonstrate improved quality of seabird distribution models by incorporating individual tracking data
Method: Locations from tracked sooty shearwaters have been aggregated into kernel density estimates. These utilization distributions will be used in conjunction with more commonly used variables to create distribution models. The model of choice is GAMLSS, for its ability to handle non-linear relationships and errors.

1. Load packages
```{r load_packages, message = FALSE}
library(dplyr)
library(gamlss)
library(foreach)
```

2. Load data
```{r load_data}
sosh.data <- read.csv('TelemetryTransect17May2016_SOSH-PFSH-COMU.csv') %>% 
  dplyr::select(SOSHcount, Binarea, Month, 
                Latitude, DistCoast, Dist200, DepCI,
                SSTmean, STD_SST, MEAN_Beaufort, L10CHLproxy, STD_CHL_log10, Watermass,
                L10CHLsurvey, CHLsurvanom, L10CHLsurvclim, FCPI) %>%
  filter(Month > 2) %>% # exclude tiny January, February surveys
  mutate(SOSHcount = as.integer(SOSHcount),
         Month = factor(Month)) %>%
  na.omit
head(sosh.data)
summary(sosh.data)
```

3. Create geographic and oceanographic models by month using canonical discrete distributions WITHOUT home range variable:
```{r fit_models, cache = TRUE}
# Create models using canonical discrete distributions.
# SOSHcount as function of FCPI, Dist200, SSTmean (all three cubic splines) with Month as random variable 
# and log(Binarea) as offset
# Create models using canonical discrete distributions for each month.
# Geographic + oceanographic models
# geo = SOSHcount ~  cs(Latitude)+cs(DistCoast)+cs(Dist200)+cs(DepCI) + offset(log(Binarea))
# ocean = SOSHcount ~ cs(SSTmean)+cs(STD_SST)+cs(MEAN_Beaufort)+cs(L10CHLproxy)+cs(STD_CHL_log10)+Watermass+cs(L10CHLsurvey)+
#   cs(CHLsurvanom)+cs(L10CHLsurvclim)+cs(FCPI) + offset(log(Binarea))
discrete.dist <- c('PO', 'NBI', 'NBII', 'DEL', 'PIG', 'SI', 'SICHEL', 'ZIP', 'ZIP2')
months <- c(6, 7, 9, 10)
# For each discrete distribution...
monthly.geo.ocean.models <- foreach(dist = discrete.dist) %do% {
  # For each month's survey...
  foreach(month = months) %do% {
    print(sprintf('Distribution %s, month %i', dist, month))
    
    # Filter SOSH data to the month of interest
    month.data <- filter(sosh.data, Month == month)
    if(nrow(month.data) == 0) stop(sprintf('No data in month %i', month))
    
    # Fit geographic model
    print('Fitting geographic model')
    geo.model <- try(gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)),
                            data = month.data,
                            family = get('dist')))
    # Try one round of refitting if it doesn't converge
    if(class(geo.model) != 'try-error' && !getElement(geo.model, 'converged')) {
      print('Refitting geographic model')
      try(geo.model <- refit(geo.model))
    }
    
    # Fit oceanographic model
    print('Fitting oceanographic model')
    ocean.model <- try(gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                                cs(STD_CHL_log10) + Watermass + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                                cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)),
                              data = month.data,
                              family = get('dist')))
    # Try one round of refitting if it doesn't converge
    if(class(ocean.model) != 'try-error' && !getElement(ocean.model, 'converged')) {
      print('Refitting oceanographic model')
      try(ocean.model <- refit(ocean.model))
    }
    
    list(dist = dist,
         month = month,
         geo.model = geo.model,
         ocean.model = ocean.model)
  }
}
```

4. Summarize model results for ranking (parameters, convergence, EDF, AIC)
```{r summarize_models}
model.rankings <- foreach(dist.list = monthly.geo.ocean.models, .combine = rbind) %do% {
  foreach(models = dist.list, .combine = rbind) %do% {
    data.frame(Distribution = models$dist,
               Month = models$month,
               GeoConverged = tryCatch(getElement(models$geo.model, 'converged'), error = function(e) NA),
               GeoEDF = tryCatch(extractAIC(models$geo.model)[1], error = function(e) NA),
               GeoAIC = tryCatch(extractAIC(models$geo.model)[2], error = function(e) NA),
               OceanConverged = tryCatch(getElement(models$ocean.model, 'converged'), error = function(e) NA),
               OceanEDF = tryCatch(extractAIC(models$ocean.model)[1], error = function(e) NA),
               OceanAIC = tryCatch(extractAIC(models$ocean.model)[2], error = function(e) NA))
  }
}
model.rankings
```

5. The best fitting models mostly come from the July survey. The SI and SICHEL distributions worked best for both types of models.
```{r best_fits}
arrange(model.rankings, -GeoConverged, GeoAIC) %>% slice(1:5)
arrange(model.rankings, -OceanConverged, OceanAIC) %>% slice(1:5)
```
