---
title: "SOSH GAMLSS"
author: "Max Czapanskiy"
date: "August 18, 2016"
output: 
  html_document: 
    toc: yes
---

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 120)
```

Goal: demonstrate improved quality of seabird distribution models by incorporating individual tracking data
Method: Locations from tracked sooty shearwaters have been aggregated into kernel density estimates. These utilization distributions will be used in conjunction with more commonly used variables to create distribution models. The model of choice is GAMLSS, for its ability to handle non-linear relationships and errors.

1. Load packages
```{r load_packages, message = FALSE, warning = FALSE}
library(gamlss)
library(dplyr)
library(foreach)
library(doParallel)
library(MASS)
library(ggplot2)
library(tidyr)
```

2. Load data
```{r load_data}
# Read data
sosh.data <- read.csv('TelemetryTransect17May2016_SOSH-PFSH-COMU.csv') %>% 
  dplyr::select(SOSHcount, Binarea, Month, 
                Latitude, DistCoast, Dist200, DepCI,
                SSTmean, STD_SST, MEAN_Beaufort, L10CHLproxy, STD_CHL_log10, Watermass,
                L10CHLsurvey, CHLsurvanom, L10CHLsurvclim, FCPI,
                SOSH0610UD, SOSH0607UD = sosh0607MN, SOSH0910UD = sosh0910MN) %>%
  filter(Month > 2) %>% # exclude tiny January, February surveys
  mutate(SOSHcount = as.integer(SOSHcount),
         Month = factor(Month),
         SOSH0607UD = ifelse(is.na(SOSH0607UD) && Month %in% 9:10, 0, SOSH0607UD),
         SOSH0910UD = ifelse(is.na(SOSH0910UD) && Month %in% 6:7, 0, SOSH0910UD)) %>% 
  na.omit
head(sosh.data)
summary(sosh.data)
```

3. Set up parallel processing
```{r register_parallel}
nCores <- detectCores()
gamlssCl <- makeCluster(nCores)
registerDoParallel(gamlssCl)
```

4. Create geographic and oceanographic models by month using canonical discrete distributions WITHOUT home range variable:
```{r fit_models, message=FALSE, warning=FALSE, cache=TRUE}
# Wrapper for gamlss. Error handling and refitting.
N_REFIT <- 5
try.fit <- function(formula, data, family, n_refit = N_REFIT) {
  # Try to fit model
  model <- try(gamlss(formula, data = data, family = family))
  fitAttempts <- 1
  
  # Keep refitting model until...
  while(fitAttempts <= n_refit &&             # ... we run out of attempts,
        class(model) != 'try-error' &&        # get an error,
        !getElement(model, 'converged')) {    # or model converges
    model <- try(refit(model))
    fitAttempts <- fitAttempts + 1
  }
  # Return fitting results. Hopefully it's a fitted model, may be a try-error
  model
}

# Create models using canonical discrete distributions.
# SOSHcount as function of FCPI, Dist200, SSTmean (all three cubic splines) with Month as random variable 
# and log(Binarea) as offset
# Create models using canonical discrete distributions for each month.
# Geographic + oceanographic models
# geo = SOSHcount ~  cs(Latitude)+cs(DistCoast)+cs(Dist200)+cs(DepCI) + offset(log(Binarea))
# ocean = SOSHcount ~ cs(SSTmean)+cs(STD_SST)+cs(MEAN_Beaufort)+cs(L10CHLproxy)+cs(STD_CHL_log10)+Watermass+cs(L10CHLsurvey)+
#   cs(CHLsurvanom)+cs(L10CHLsurvclim)+cs(FCPI) + offset(log(Binarea))
discrete.dist <- c('PO', 'NBI', 'NBII', 'DEL', 'PIG', 'SI', 'SICHEL', 'ZIP', 'ZIP2')
months <- c(6, 7, 9, 10)
# For each discrete distribution...
monthly.geo.ocean.models <- foreach(dist = discrete.dist,
                                    .packages = c('foreach',
                                                  'doParallel',
                                                  'gamlss',
                                                  'dplyr')) %dopar% {
  # For each month's survey...
  foreach(month = months) %do% {
    print(sprintf('Distribution %s, month %i', dist, month))
    
    # Filter SOSH data to the month of interest
    month.data <- filter(sosh.data, Month == month)
    if(nrow(month.data) == 0) stop(sprintf('No data in month %i', month))
    
    # Fit geographic model
    print('Fitting geographic model')
    geo.model <- try.fit(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)),
                         data = month.data,
                         family = get('dist'))
    
    # Fit oceanographic model
    print('Fitting oceanographic model')
    ocean.model <- try.fit(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                                cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                                cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)),
                           data = month.data,
                           family = get('dist'))
    
    list(dist = dist,
         month = month,
         geo.model = geo.model,
         ocean.model = ocean.model)
  }
}

stopCluster(gamlssCl)
```

5. Summarize model results for ranking (parameters, convergence, EDF, AIC)
```{r summarize_models}
model.rankings <- foreach(dist.list = monthly.geo.ocean.models, .combine = rbind) %do% {
  foreach(models = dist.list, .combine = rbind) %do% {
    data.frame(Distribution = models$dist,
               Month = models$month,
               GeoConverged = tryCatch(getElement(models$geo.model, 'converged'), error = function(e) NA),
               GeoEDF = tryCatch(extractAIC(models$geo.model)[1], error = function(e) NA),
               GeoAIC = tryCatch(extractAIC(models$geo.model)[2], error = function(e) NA),
               OceanConverged = tryCatch(getElement(models$ocean.model, 'converged'), error = function(e) NA),
               OceanEDF = tryCatch(extractAIC(models$ocean.model)[1], error = function(e) NA),
               OceanAIC = tryCatch(extractAIC(models$ocean.model)[2], error = function(e) NA))
  }
}
model.rankings
```

6. SI, SICHEL, and PIG distributions perform the best (which makes sense, they're all closely related), but only SI and SICHEL are in each month's top 4 distributions for both geographic and oceanographic models. SI has the slightly greater mean AIC weight than SICHEL, so we'll proceed using SI.
```{r best_fits}
# Utility function for calculating AIC weight. vAIC is a vector of AIC values
calcAICw <- function(vAIC) {
  deltaAIC <- vAIC - min(vAIC, na.rm = TRUE)
  relLikelihood <- exp(-0.5 * deltaAIC)
  normalizingFactor <- sum(relLikelihood, na.rm = TRUE)
  relLikelihood / normalizingFactor
}

# Sort models by month and quality of fit
monthly.geo.ranks <- model.rankings %>%
  group_by(Month) %>%
  mutate(GeoDeltaAIC = GeoAIC - min(GeoAIC, na.rm = TRUE),
         GeoAICw = calcAICw(GeoAIC) %>% round(3)) %>%
  arrange(-GeoAICw) %>% 
  slice(1:4) %>% 
  ungroup %>%
  select(Distribution:GeoAICw)
monthly.ocean.ranks <- model.rankings %>%
  group_by(Month) %>%
  mutate(OceanDeltaAIC = OceanAIC - min(OceanAIC, na.rm = TRUE),
         OceanAICw = calcAICw(OceanAIC) %>% round(3)) %>%
  arrange(-OceanAICw) %>% 
  slice(1:4) %>%
  ungroup %>%
  select(Distribution, Month, OceanConverged:OceanAICw)

# Identify best available distribution by product of AIC weight across months
monthly.geo.ranks %>% 
  group_by(Distribution) %>%
  summarize(N = n(),
            meanAICw = mean(GeoAICw)) %>%
  ungroup %>%
  arrange(-N, -meanAICw)

monthly.ocean.ranks %>% 
  group_by(Distribution) %>%
  summarize(N = n(),
            meanAICw = mean(OceanAICw)) %>%
  ungroup %>%
  arrange(-N, -meanAICw)
```

7. Model culling
```{r cull_models, warning=FALSE, cache=TRUE, results='hide'}
# We see SI and SICHEL are the only two distributions to show up in both models' top 4 across months
# SI has the slightly higher mean AIC weight, so we'll go with that.

# Cull models using selected distribution (SI)

# Recursively drop parameters by largest decrease in AIC until model degrades
cull.model <- function(model) {
  aic <- extractAIC(model)[2]
    
  # Re-fit model to n-1 terms
  model.terms <- model %>% formula %>% terms
  term.labels <- attr(model.terms, 'term.labels')
  submodels <- foreach(i = seq(term.labels)) %do% {
    dropped <- term.labels[i]
    submodel <- try(update(model, reformulate(sprintf('. - %s', dropped))))
    list(dropped = dropped,
         submodel = submodel)
  }
  
  # Which submodels converged? Which submodel has the greatest decrease in AIC?
  drop.results <- foreach(i = seq(submodels), .combine = rbind) %do% {
    tryCatch(with(submodels[[i]], data.frame(dropped = dropped, 
                                             i = i,
                                             converged = submodel$converged,
                                             AIC = extractAIC(submodel)[2],
                                             deltaAIC = extractAIC(submodel)[2] - aic)),
             error = function(e) data.frame(dropped = NA,
                                            i = NA,
                                            converged = NA,
                                            AIC = NA,
                                            deltaAIC = NA))
  } %>%
    filter(converged) %>%
    arrange(deltaAIC)
  
  # If no submodels converge or if no submodels are an improvement, return the original model
  if(nrow(drop.results) == 0 || min(drop.results$deltaAIC, na.rm = TRUE) > 0) {
    return(model)
  } else {
    # Otherwise, repeat the process on the best submodel
    best.submodel <- submodels[[drop.results$i[1]]]$submodel
    return(cull.model(best.submodel))
  }
}

# Month 6 
month6 <- filter(sosh.data, Month == 6)

# Geo
geo6global <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)),
                     data = month6,
                     family = SI,
                     control = gamlss.control(n.cyc = 100))
geo6dropterm <- dropterm(geo6global, test = 'Chisq')
geo6culled <- cull.model(geo6global)
geo6dropped <- setdiff(terms(geo6global) %>% attr('term.labels'), terms(geo6culled) %>% attr('term.labels'))

# Ocean
ocean6global <- gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                         cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                         cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)),
                       data = month6,
                       family = SI,
                       control = gamlss.control(n.cyc = 100))
ocean6dropterm <- dropterm(ocean6global, test = 'Chisq')
ocean6culled <- cull.model(ocean6global)
ocean6dropped <- setdiff(terms(ocean6global) %>% attr('term.labels'), terms(ocean6culled) %>% attr('term.labels'))

# Month 7 
month7 <- filter(sosh.data, Month == 7)

# Geo
geo7global <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)),
                     data = month7,
                     family = SI,
                     control = gamlss.control(n.cyc = 100))
geo7dropterm <- dropterm(geo7global, test = 'Chisq')
geo7culled <- cull.model(geo7global)
geo7dropped <- setdiff(terms(geo7global) %>% attr('term.labels'), terms(geo7culled) %>% attr('term.labels'))

# Ocean
ocean7global <- try(gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                             cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                             cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)),
                           data = month7,
                           family = SI,
                           control = gamlss.control(n.cyc = 100)))
ocean7dropterm <- dropterm(ocean7global, test = 'Chisq')
ocean7culled <- cull.model(ocean7global)
ocean7dropped <- setdiff(terms(ocean7global) %>% attr('term.labels'), terms(ocean7culled) %>% attr('term.labels'))

# Month 9 
month9 <- filter(sosh.data, Month == 9)

# Geo
# For some reason month 9 geo model doesn't converge reliably
geo9global <- try.fit(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)),
                      data = month9,
                      family = SI)
geo9dropterm <- dropterm(geo9global, test = 'Chisq')
geo9culled <- cull.model(geo9global)
geo9dropped <- setdiff(terms(geo9global) %>% attr('term.labels'), terms(geo9culled) %>% attr('term.labels'))

# Ocean
ocean9global <- gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                         cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                         cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)),
                       data = month9,
                       family = SI,
                       control = gamlss.control(n.cyc = 100))
ocean9dropterm <- dropterm(ocean9global, test = 'Chisq')
ocean9culled <- cull.model(ocean9global)
ocean9dropped <- setdiff(terms(ocean9global) %>% attr('term.labels'), terms(ocean9culled) %>% attr('term.labels'))

# Month 10 
month10 <- filter(sosh.data, Month == 10)

# Geo
geo10global <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)),
                     data = month10,
                     family = SI,
                     control = gamlss.control(n.cyc = 100))
geo10dropterm <- dropterm(geo10global, test = 'Chisq')
geo10culled <- cull.model(geo10global)
geo10dropped <- setdiff(terms(geo10global) %>% attr('term.labels'), terms(geo10culled) %>% attr('term.labels'))

# Ocean
ocean10global <- try(gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                             cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) +
                             cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)),
                           data = month10,
                           family = SI,
                           control = gamlss.control(n.cyc = 100)))
ocean10dropterm <- dropterm(ocean10global, test = 'Chisq')
ocean10culled <- cull.model(ocean10global)
ocean10dropped <- setdiff(terms(ocean10global) %>% attr('term.labels'), terms(ocean10culled) %>% attr('term.labels'))
```

8. Per-month combined models
```{r combine_models, cache = TRUE, results = 'hide'}
combined6culled <- gamlss(formula = SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) +
                      cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                      as.factor(Watermass) + cs(L10CHLsurvey) + cs(L10CHLsurvclim) + 
                      cs(FCPI) + offset(log(Binarea)), 
                    family = SI, 
                    data = month6,  
                    control = gamlss.control(n.cyc = 100)) 

combined7culled <- gamlss(formula = SOSHcount ~ cs(Latitude) + cs(DistCoast) + 
                      cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                    family = SI, 
                    data = month7,  
                    control = gamlss.control(n.cyc = 100)) 

# Combined model fails for month 9
# combined9culled <- gamlss(formula = SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) +
#                       cs(SSTmean) + cs(MEAN_Beaufort) +  
#                       cs(STD_CHL_log10) + cs(CHLsurvanom) + cs(L10CHLsurvclim) +  
#                       cs(FCPI) + offset(log(Binarea)), 
#                     family = SI, 
#                     data = month9,  
#                     control = gamlss.control(n.cyc = 100)) 

combined10culled <- gamlss(formula = SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + 
                       cs(SSTmean) + cs(STD_SST) + cs(L10CHLproxy) +  
                       as.factor(Watermass) + cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                    family = SI, 
                    data = month10,  
                    control = gamlss.control(n.cyc = 100)) 
```

9. Model analysis
```{r model_analysis}
model.options <- foreach(type = c('geo', 'ocean', 'combined'), .combine = rbind) %do% {
  foreach(month = c(6, 7, 9, 10), .combine = rbind) %do% {
    foreach(scope = c('global', 'culled'), .combine = rbind) %do% {
      if(type == 'combined' && scope == 'global')
        return(NULL)
      model <- try(get(paste(type, month, scope, sep = '')))
      if(class(model)[1] == 'try-error')
        return(NULL)
      aic <- extractAIC(model)[2]
      predictors <- model %>% formula %>% terms %>% attr('term.labels') %>% paste(collapse = ', ')
      data.frame(type = type,
                 month = month,
                 scope = scope,
                 AIC = aic,
                 predictors = predictors)
    }
  }
} %>%
  group_by(month) %>%
  mutate(AICw = calcAICw(AIC)) %>%
  ungroup %>%
  arrange(month, type, scope)

model.options

ggplot(model.options,
       aes(x = paste(type, scope),
           y = AICw)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~ month,
             nrow = 4) +
  scale_x_discrete('Model Option',
                   limits = c('geo global', 'geo culled',
                              'ocean global', 'ocean culled',
                              'combined culled')) +
  ggtitle('Relative Model Performance')

predictor.comparison <- foreach(type = c('geo', 'ocean', 'combined'), .combine = rbind) %do% {
  foreach(month = c(6, 7, 9, 10), .combine = rbind) %do% {
    foreach(scope = c('global', 'culled'), .combine = rbind) %do% {
      if(type == 'combined' && scope == 'global')
        return(NULL)
      model <- try(get(paste(type, month, scope, sep = '')))
      if(class(model)[1] == 'try-error')
        return(NULL)
      predictors <- model %>% formula %>% terms %>% attr('term.labels')
      data.frame(type = type,
                 month = month,
                 scope = scope,
                 predictor = predictors)
    }
  }
}

geo.predictor.prevalence <- predictor.comparison %>% 
  filter(type == 'geo',
         scope == 'culled') %>%
  group_by(predictor) %>%
  summarize(N = n()) %>%
  ungroup() %>%
  arrange(N, predictor)

predictor.comparison %>%
  filter(type == 'geo', scope == 'culled') %>%
  mutate(predictor = factor(predictor, levels = geo.predictor.prevalence$predictor)) %>%
  ggplot(aes(x = factor(month), 
             y = predictor)) +
  geom_tile(fill = '#045a8d') +
  labs(title = 'Predictor Retention\nGeographic Models',
       x = 'Month',
       y = 'Predictor') 

ocean.predictor.prevalence <- predictor.comparison %>% 
  filter(type == 'ocean',
         scope == 'culled') %>%
  group_by(predictor) %>%
  summarize(N = n()) %>%
  ungroup() %>%
  arrange(N, predictor)

predictor.comparison %>%
  filter(type == 'ocean', scope == 'culled') %>%
  mutate(predictor = factor(predictor, levels = ocean.predictor.prevalence$predictor)) %>%
  ggplot(aes(x = factor(month), 
             y = predictor)) +
  geom_tile(fill = '#006d2c') +
  labs(title = 'Predictor Retention\nOceanographic Models',
       x = 'Month',
       y = 'Predictor') 
```

10. Add UD variables
```{r ud_models, cache = TRUE, results = 'hide'}
# Adding in UD variable

# Code generator
# The commented out section creates the code below it

# foreach(month = c(6:7, 9:10), .combine = c) %do% {
#   foreach(type = c('geo', 'ocean', 'combined'), .combine = c) %do% {
#     foreach(scope = c('global', 'culled'), .combine = c) %do% {
#       foreach(ud = c('all', 'season'), .combine = c) %do% {
#         if(type == 'combined' && scope == 'global')
#           return(NULL)
#         
#         varname <- if(ud == 'all') {
#           '0610'
#         } else if(month %in% 6:7) {
#           '0607'
#         } else {
#           '0910'
#         }
#         
#         sprintf('%s%i%sUD%s <- tryCatch(update(%s%i%s, ~ . + SOSH%sUD), error = function(e) NA)', type, month, scope, ud, type, month, scope, varname)
#       }
#     }
#   }
# } %>% paste(collapse = '\n') %>% cat

# Month 6
geo6globalUDall <- tryCatch(update(geo6global, ~ . + SOSH0610UD), error = function(e) NA)
geo6globalUDseason <- tryCatch(update(geo6global, ~ . + SOSH0607UD), error = function(e) NA)
geo6culledUDall <- tryCatch(update(geo6culled, ~ . + SOSH0610UD), error = function(e) NA)
geo6culledUDseason <- tryCatch(update(geo6culled, ~ . + SOSH0607UD), error = function(e) NA)
ocean6globalUDall <- tryCatch(update(ocean6global, ~ . + SOSH0610UD), error = function(e) NA)
ocean6globalUDseason <- tryCatch(update(ocean6global, ~ . + SOSH0607UD), error = function(e) NA)
ocean6culledUDall <- tryCatch(update(ocean6culled, ~ . + SOSH0610UD), error = function(e) NA)
ocean6culledUDseason <- tryCatch(update(ocean6culled, ~ . + SOSH0607UD), error = function(e) NA)
combined6culledUDall <- tryCatch(update(combined6culled, ~ . + SOSH0610UD), error = function(e) NA)
combined6culledUDseason <- tryCatch(update(combined6culled, ~ . + SOSH0607UD), error = function(e) NA)

# Month 7
geo7globalUDall <- tryCatch(update(geo7global, ~ . + SOSH0610UD), error = function(e) NA)
geo7globalUDseason <- tryCatch(update(geo7global, ~ . + SOSH0607UD), error = function(e) NA)
geo7culledUDall <- tryCatch(update(geo7culled, ~ . + SOSH0610UD), error = function(e) NA)
geo7culledUDseason <- tryCatch(update(geo7culled, ~ . + SOSH0607UD), error = function(e) NA)
ocean7globalUDall <- tryCatch(update(ocean7global, ~ . + SOSH0610UD), error = function(e) NA)
ocean7globalUDseason <- tryCatch(update(ocean7global, ~ . + SOSH0607UD), error = function(e) NA)
ocean7culledUDall <- tryCatch(update(ocean7culled, ~ . + SOSH0610UD), error = function(e) NA)
ocean7culledUDseason <- tryCatch(update(ocean7culled, ~ . + SOSH0607UD), error = function(e) NA)
combined7culledUDall <- tryCatch(update(combined7culled, ~ . + SOSH0610UD), error = function(e) NA)
combined7culledUDseason <- tryCatch(update(combined7culled, ~ . + SOSH0607UD), error = function(e) NA)

# Month 9
geo9globalUDall <- tryCatch(update(geo9global, ~ . + SOSH0610UD), error = function(e) NA)
geo9globalUDseason <- tryCatch(update(geo9global, ~ . + SOSH0910UD), error = function(e) NA)
geo9culledUDall <- tryCatch(update(geo9culled, ~ . + SOSH0610UD), error = function(e) NA)
geo9culledUDseason <- tryCatch(update(geo9culled, ~ . + SOSH0910UD), error = function(e) NA)
ocean9globalUDall <- tryCatch(update(ocean9global, ~ . + SOSH0610UD), error = function(e) NA)
ocean9globalUDseason <- tryCatch(update(ocean9global, ~ . + SOSH0910UD), error = function(e) NA)
ocean9culledUDall <- tryCatch(update(ocean9culled, ~ . + SOSH0610UD), error = function(e) NA)
ocean9culledUDseason <- tryCatch(update(ocean9culled, ~ . + SOSH0910UD), error = function(e) NA)
combined9culledUDall <- tryCatch(update(combined9culled, ~ . + SOSH0610UD), error = function(e) NA)
combined9culledUDseason <- tryCatch(update(combined9culled, ~ . + SOSH0910UD), error = function(e) NA)

# Month 10
geo10globalUDall <- tryCatch(update(geo10global, ~ . + SOSH0610UD), error = function(e) NA)
geo10globalUDseason <- tryCatch(update(geo10global, ~ . + SOSH0910UD), error = function(e) NA)
geo10culledUDall <- tryCatch(update(geo10culled, ~ . + SOSH0610UD), error = function(e) NA)
geo10culledUDseason <- tryCatch(update(geo10culled, ~ . + SOSH0910UD), error = function(e) NA)
ocean10globalUDall <- tryCatch(update(ocean10global, ~ . + SOSH0610UD), error = function(e) NA)
ocean10globalUDseason <- tryCatch(update(ocean10global, ~ . + SOSH0910UD), error = function(e) NA)
ocean10culledUDall <- tryCatch(update(ocean10culled, ~ . + SOSH0610UD), error = function(e) NA)
ocean10culledUDseason <- tryCatch(update(ocean10culled, ~ . + SOSH0910UD), error = function(e) NA)
combined10culledUDall <- tryCatch(update(combined10culled, ~ . + SOSH0610UD), error = function(e) NA)
combined10culledUDseason <- tryCatch(update(combined10culled, ~ . + SOSH0910UD), error = function(e) NA)
```

11. UD Model analysis
```{r ud_model_analysis}
# Check UD models against base
UDcomparison <- foreach(month = c(6:7, 9:10), .combine = rbind) %do% {
  foreach(type = c('geo', 'ocean', 'combined'), .combine = rbind) %do% {
    foreach(scope = c('global', 'culled'), .combine = rbind) %do% {
      if(type == 'combined' && scope == 'global')
        return(NULL)
      
      tryCatch({
        model.name <- sprintf('%s%i%s', type, month, scope) 
        base <- get(model.name)
        udall <- get(sprintf('%s%s', model.name, 'UDall'))
        udseason <- get(sprintf('%s%s', model.name, 'UDseason'))
        
        data.frame(model = model.name,
                   UD = c('base', 'UDall', 'UDseason'),
                   AIC = c(extractAIC(base)[2],
                           extractAIC(udall)[2],
                           extractAIC(udseason)[2])) %>%
          mutate(AICw = calcAICw(AIC))
      }, error = function(e) NULL)
    }
  }
}

# View summary of results and plot
UDcomparison %>% 
  select(model, UD, AICw) %>% 
  spread(UD, AICw)

ggplot(UDcomparison, 
       aes(x = model,
           y = AICw, 
           fill = UD)) + 
  geom_bar(stat = 'identity') +
  scale_fill_discrete('UD Usage', labels = c('None', 'All\n(6-10)', 'Season\n(6-7, 9-10)')) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = 'Relative Performance of UD Models',
       x = 'Base Model')
```
