---
title: "SOSH Distribution Model + UD"
author: "Max Czapanskiy"
date: "September 13, 2016"
output: html_document
---

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 120)
```

## Hypothesis
Telemetry data can be used to enhance fitness of distribution models by including utilization distributions derived from individual tracks as a predictor.

## Approach
Create traditional distribution models using environmental variables to predict at-sea counts. Add utilization distributions (UDs) and test if fit quality improved.

## Method
### Choose response variable distribution

1. We divided our count data into four survey months. 
2. For each month, we created two models (one using geographic predictors, the other oceanographic) using canonical count distributions: Poisson, Poisson inverse Gaussian, zero-inflated Poisson, negative binomial, Delacorte, and Sichel. 
3. We found that models using the Sichel distribution converged most frequently with the lowest AIC.

### Identify best model for each month

1. To maximize goodness of fit while reducing model complexity, we used iterative parameter removal and gAIC to discard parameters in each monthly global geographic (4 parameters) model and each monthly global oceanographic (10 parameters) model that did not show appreciable change in model fit (e.g., if removal of a parameter did not change AIC or if AIC increased with a parameter’s removal, it was dropped). 
2. Resulting “culled geographic”, and “culled oceanographic” models were then combined, and the culling process was repeated[AJ1] . 
3. The final, “best” model for each month then was determined using model ranking based on gAIC. The parameter culling process allowed us to reduce 14 parameters (plus the offset) to a maximum of 3 – 7 explanatory parameters, thereby reducing the overall complexity of the models. 
4. The resulting “combined culled” model was selected as the best explanatory model in 3 of 4 months, but in month 7, the “geo culled” also was retained as a competing model (Δ gAIC < 2). In month 9, the global geo model was >35-times more likely, based on gAIC, to be selected than the next highest ranking model. 

```{r load_pkgs_data, message = FALSE, warning = FALSE}
library(gamlss)
library(dplyr)
library(foreach)
library(iterators)
library(knitr)

sosh.data <- read.csv('TelemetryTransect17May2016_SOSH-PFSH-COMU.csv') %>% 
  dplyr::select(SOSHcount, Binarea, Month, 
                Latitude, DistCoast, Dist200, DepCI,
                SSTmean, STD_SST, MEAN_Beaufort, L10CHLproxy, STD_CHL_log10, Watermass,
                L10CHLsurvey, CHLsurvanom, L10CHLsurvclim, FCPI,
                SOSH0610UD, SOSH0607UD = sosh0607MN, SOSH0910UD = sosh0910MN) %>%
  filter(Month > 2) %>% # exclude tiny January, February surveys
  mutate(SOSHcount = as.integer(SOSHcount),
         Month = factor(Month),
         # Replace NAs with 0s in UD columns
         SOSH0607UD = ifelse(is.na(SOSH0607UD) && Month %in% 9:10, 0, SOSH0607UD),
         SOSH0910UD = ifelse(is.na(SOSH0910UD) && Month %in% 6:7, 0, SOSH0910UD)) %>% 
  na.omit
```

```{r base_models_6, message = FALSE, warning = FALSE, cache = TRUE}
month6 <- filter(sosh.data, Month == 6)

geo6global <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)), 
                    family = SI, 
                    data = month6,  
                    control = gamlss.control(n.cyc = 100))

geo6culled <- geo6global

ocean6global <- gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                         cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                         cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                    family = SI, 
                    data = month6,  
                    control = gamlss.control(n.cyc = 100))

ocean6culled <- gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                         as.factor(Watermass) + cs(L10CHLsurvey) + cs(L10CHLsurvclim) + 
                         cs(FCPI) + offset(log(Binarea)), 
                    family = SI, 
                    data = month6,  
                    control = gamlss.control(n.cyc = 100))

combined6culled <- gamlss(SOSHcount ~ cs(Latitude) + cs(Dist200) + cs(SSTmean) + cs(STD_SST) + 
                            cs(MEAN_Beaufort) + cs(L10CHLproxy) + cs(L10CHLsurvey) + offset(log(Binarea)), 
                    family = SI, 
                    data = month6,  
                    control = gamlss.control(n.cyc = 100))
```

```{r base_models_7, message = FALSE, warning = FALSE, cache = TRUE}
month7 <- filter(sosh.data, Month == 7)

geo7global <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)), 
                    family = SI, 
                    data = month7,  
                    control = gamlss.control(n.cyc = 100))

geo7culled <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + offset(log(Binarea)), 
                    family = SI, 
                    data = month7,  
                    control = gamlss.control(n.cyc = 100))

ocean7global <- gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                         cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                         cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                       family = SI, 
                       data = month7,  
                       control = gamlss.control(n.cyc = 100))

ocean7culled <- gamlss(SOSHcount ~ cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                       family = SI, 
                       data = month7,  
                       control = gamlss.control(n.cyc = 100))

combined7culled <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(FCPI) + offset(log(Binarea)), 
                          family = SI, 
                          data = month7,  
                          control = gamlss.control(n.cyc = 100))
```

```{r base_models_9, message = FALSE, warning = FALSE, cache = TRUE}
month9 <- filter(sosh.data, Month == 9)

# Gamlss fitting for month 9 is buggy. Re-fitting works, n.cyc = 100 does not.
N_REFIT <- 5
try.fit <- function(formula, data, family, n_refit = N_REFIT) {
  # Try to fit model
  model <- try(gamlss(formula, data = data, family = family))
  fitAttempts <- 1
  
  # Keep refitting model until...
  while(fitAttempts <= n_refit &&             # ... we run out of attempts,
        class(model) != 'try-error' &&        # get an error,
        !getElement(model, 'converged')) {    # or model converges
    model <- try(refit(model))
    fitAttempts <- fitAttempts + 1
  }
  # Return fitting results. Hopefully it's a fitted model, may be a try-error
  model
}

geo9global <- try.fit(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)), 
                      family = SI, 
                      data = month9)

geo9culled <- geo9global

ocean9global <- tryCatch(gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                                  cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                                  cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                                family = SI, 
                                data = month9,  
                                control = gamlss.control(n.cyc = 100)),
                         error = function(e) NULL)

ocean9culled <- tryCatch(gamlss(SOSHcount ~ cs(SSTmean) + cs(MEAN_Beaufort) + cs(STD_CHL_log10) + cs(CHLsurvanom) + 
                                  cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                                family = SI, 
                                data = month9,  
                                control = gamlss.control(n.cyc = 100)),
                         error = function(e) NULL)

combined9culled <- NULL
```

```{r base_models_10, message = FALSE, warning = FALSE, cache = TRUE}
month10 <- filter(sosh.data, Month == 10)

geo10global <- tryCatch(gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + offset(log(Binarea)), 
                               family = SI, 
                               data = month10,  
                               control = gamlss.control(n.cyc = 100)),
                        error = function(e) NULL)

geo10culled <- tryCatch(gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + offset(log(Binarea)), 
                               family = SI, 
                               data = month10,  
                               control = gamlss.control(n.cyc = 100)),
                        error = function(e) NULL)

ocean10global <- tryCatch(gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(MEAN_Beaufort) + cs(L10CHLproxy) + 
                                   cs(STD_CHL_log10) + as.factor(Watermass) + cs(L10CHLsurvey) + cs(CHLsurvanom) + 
                                   cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                                 family = SI, 
                                 data = month10,  
                                 control = gamlss.control(n.cyc = 100)),
                          error = function(e) NULL)

ocean10culled <- tryCatch(gamlss(SOSHcount ~ cs(SSTmean) + cs(STD_SST) + cs(L10CHLproxy) + as.factor(Watermass) + 
                                   cs(L10CHLsurvclim) + cs(FCPI) + offset(log(Binarea)), 
                                 family = SI, 
                                 data = month10,  
                                 control = gamlss.control(n.cyc = 100)),
                          error = function(e) NULL)

combined10culled <- tryCatch(gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(SSTmean) + cs(STD_SST) + 
                                      cs(L10CHLproxy) + cs(L10CHLsurvclim) + offset(log(Binarea)), 
                                    family = SI, 
                                    data = month10,  
                                    control = gamlss.control(n.cyc = 100)),
                             error = function(e) NULL)
```

```{r base_summaries, cache = TRUE}
formula.to.char <- function(model) {
  model %>% formula %>% terms.formula %>% attr('term.labels') %>% paste(collapse = ' + ')
}

# Utility function for calculating AIC weight. vAIC is a vector of AIC values
calcAICw <- function(vAIC) {
  deltaAIC <- vAIC - min(vAIC, na.rm = TRUE)
  relLikelihood <- exp(-0.5 * deltaAIC)
  normalizingFactor <- sum(relLikelihood, na.rm = TRUE)
  relLikelihood / normalizingFactor
}

# For all combinations of month, type, and scope...
model.summary.0 <- expand.grid(Month = c(6, 7, 9, 10),
                             type = c('geo', 'ocean', 'combined'),
                             scope = c('global', 'culled'),
                             stringsAsFactors = FALSE) %>%
  filter(!(type == 'combined' & scope == 'global')) %>% # There are no combined global models 
  mutate(Model = paste(type, scope)) %>%
  arrange(Month, type, scope)

model.summary <- foreach(row = iter(model.summary.0, by = 'row'), .combine = rbind) %do% {
  with(row, {
    # Fetch model
    model <- try(get(sprintf('%s%i%s', type, Month, scope)))

    # Verify model exists
    if(class(model)[1] == 'try-error' ||
       is.null(model))
      return(mutate(row,
                    Formula = NA,
                    gAIC = NA,
                    EDF = NA))

    # Run model diagnostics
    mutate(row,
           Formula = formula.to.char(model),
           gAIC = extractAIC(model)[2],
           EDF = extractAIC(model)[1])
  })
} %>%
  # Run model group diagnostics
  group_by(Month) %>%
  mutate(deltaGAIC = gAIC - min(gAIC, na.rm = TRUE),
         GAICw = calcAICw(gAIC)) %>%
  ungroup %>%
  select(Month, Model, Formula, gAIC, EDF, deltaGAIC, GAICw) %>%
  arrange(Month, gAIC)

```

### Base Model Comparisons
```{r base_summaries2, cache = TRUE}
kable(model.summary, digits = 2)
```

### Best Models Diagnostics
### Month 6
```{r best_diagnostics_6, cache = TRUE}
model <- combined6culled

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles", 
            plot.it = TRUE, 
            frame.plot = TRUE, 
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```

### Month 7
```{r best_diagnostics_7}
model <- combined7culled

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles", 
            plot.it = TRUE, 
            frame.plot = TRUE, 
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```

### Month 9
```{r best_diagnostics_9, cache = TRUE}
model <- geo9culled

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles",
            plot.it = TRUE,
            frame.plot = TRUE,
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
# wp(model)
# Worm plot error in month 9
 
# Term Plots
# term.plot(model)
# Term plots error in month 9
```

### Month 10
```{r best_diagnostics_10, cache = TRUE}
model <- combined10culled

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles",
            plot.it = TRUE,
            frame.plot = TRUE,
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```

### Augment monthly models with UD variable

1. Season specific (6 and 7, or 9 and 10) and full survey period (6 – 10) shearwater BBUDs were added to the top-ranking model in each month to evaluate potential for increasing model fit based on gAIC.
2. What works best?

```{r add_ud_6, cache = TRUE}
ud6seasonal <- gamlss(SOSHcount ~ cs(Latitude) + cs(Dist200) + cs(SSTmean) + cs(STD_SST) + 
                        cs(MEAN_Beaufort) + cs(L10CHLproxy) + cs(L10CHLsurvey) + offset(log(Binarea)) +
                        SOSH0607UD, 
                      family = SI, 
                      data = month6,  
                      control = gamlss.control(n.cyc = 100))
ud6total <- gamlss(SOSHcount ~ cs(Latitude) + cs(Dist200) + cs(SSTmean) + cs(STD_SST) + 
                        cs(MEAN_Beaufort) + cs(L10CHLproxy) + cs(L10CHLsurvey) + offset(log(Binarea)) +
                        SOSH0610UD, 
                      family = SI, 
                      data = month6,  
                      control = gamlss.control(n.cyc = 100))
```
```{r add_ud_7, cache = TRUE}
ud7seasonal <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(FCPI) + offset(log(Binarea)) +
                        SOSH0607UD, 
                      family = SI, 
                      data = month7,  
                      control = gamlss.control(n.cyc = 100))
ud7total <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(FCPI) + offset(log(Binarea)) +
                        SOSH0610UD, 
                      family = SI, 
                      data = month7,  
                      control = gamlss.control(n.cyc = 200))
```
```{r add_ud_9, cache = TRUE}
ud9seasonal <- try.fit(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + 
                         offset(log(Binarea)) + SOSH0910UD, 
                       family = SI, 
                       data = month9)
ud9total <- try.fit(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(Dist200) + cs(DepCI) + 
                         offset(log(Binarea)) + SOSH0610UD, 
                       family = SI, 
                       data = month9)
```
```{r add_ud_10, cache = TRUE}
ud10seasonal <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(SSTmean) + cs(STD_SST) + 
                         cs(L10CHLproxy) + cs(L10CHLsurvclim) + offset(log(Binarea)) + SOSH0910UD, 
                       family = SI, 
                       data = month10,  
                       control = gamlss.control(n.cyc = 100))
ud10total <- gamlss(SOSHcount ~ cs(Latitude) + cs(DistCoast) + cs(SSTmean) + cs(STD_SST) + 
                      cs(L10CHLproxy) + cs(L10CHLsurvclim) + offset(log(Binarea)) + SOSH0610UD, 
                    family = SI, 
                    data = month10,  
                    control = gamlss.control(n.cyc = 100))
```

```{r ud_summaries, cache = TRUE}
ud.summary.0 <- data.frame(Month = rep(c(6,7,9,10), each = 3),
                           Model = rep(c('Base', 'UD Seasonal', 'UD Total'), 4)) %>%
  arrange(Month, Model)

ud.summary <- foreach(row = iter(ud.summary.0, by = 'row'),
                      model = list(combined6culled, ud6seasonal, ud6total,
                                   combined7culled, ud7seasonal, ud7total,
                                   geo9culled, ud9seasonal, ud9total,
                                   combined10culled, ud10seasonal, ud10total),
                      .combine = rbind) %do% {
  with(row, {
    # Run model diagnostics
    mutate(row,
           Formula = formula.to.char(model),
           gAIC = extractAIC(model)[2],
           EDF = extractAIC(model)[1],
           Converged = model$converged)
  })
} %>%
  # Run model group diagnostics
  group_by(Month) %>%
  mutate(deltaGAIC = gAIC - min(gAIC, na.rm = TRUE),
         GAICw = calcAICw(gAIC)) %>%
  ungroup %>%
  select(Month, Model, Formula, gAIC, EDF, deltaGAIC, GAICw, Converged) %>%
  arrange(Month, gAIC)
```

### UD Model Comparisons
```{r ud_summaries2}
kable(ud.summary, digits = 2)
```

### UD Model Diagnostics
### Month 6
```{r ud_diagnostics_6, cache = TRUE}
model <- ud6total

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles", 
            plot.it = TRUE, 
            frame.plot = TRUE, 
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```

### Month 7
```{r ud_diagnostics_7}
model <- ud7total

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles", 
            plot.it = TRUE, 
            frame.plot = TRUE, 
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```

### Month 9
```{r ud_diagnostics_9, cache = TRUE}
model <- ud9total

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles",
            plot.it = TRUE,
            frame.plot = TRUE,
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
# wp(model)
# Worm plot error in month 9
 
# Term Plots
# term.plot(model)
# Term plots error in month 9
```

### Month 10
```{r ud_diagnostics_10, cache = TRUE}
model <- ud10seasonal

# Coefficients
summary(model)

# QQ Norm plot (Copied from plot.gamlss)
residx <- resid(model)
qqnorm(residx, main = "Normal Q-Q Plot",
            xlab = "Theoretical Quantiles",
            ylab = "Sample Quantiles",
            plot.it = TRUE,
            frame.plot = TRUE,
            col="darkgreen")
lines(residx, residx, col="red" , lwd=.4, cex=.4 )

# Worm Plot
wp(model)

# Term Plots
term.plot(model)
```
